---
title: "Visualisation-II"
author: "Revendranath T"
date: "27/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## static plot
```{r}
library(ggplot2)
iris %>% glimpse()

p <- ggplot(data = iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point()

plot(p)
#or
p
```


## ggAnimate
```{r}
# packages required
install.packages('gganimate')
install.packages('gifski')
#install.packages('png')

library(gifski)
#library(png)
library(gganimate)

class(p)

# add transition to p (ggplot) object
anim <- p + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)

anim
!```
## what changed from static to animation?
* **transition_states()** splits up plot data by a discrete variable and animates between the different states.
* Transitions are functions that interpret the plot data in order to somehow distribute it over a number of frames. 
* transition_states() specifically splits the data into subsets based on a variable in the data (here Species), and calculates intermediary data states that ensures a smooth transition between the states (something referred to as tweening). 
* gganimate provides a range of different transitions

## Easing
* Easing is to decide how the change from one value to another should progress. 
* The default easing is linear, but others can be used, potentially only targeting specific aesthetics. 
* Setting easing is done with the ease_aes() function. 
* The first argument sets the default easing and subsequent named arguments sets it for specific aesthetics.

```{r}
anim + 
  ease_aes('cubic-in-out') # Slow start and end for a smoother look
```
* ease_aes() defines the velocity with which aesthetics change during an animation.
```{r}
anim + 
  ease_aes(y = 'bounce-out') # Sets special ease for y aesthetic
anim
```
# Labeling
* It can be quite hard to understand an animation without any indication as to what each time point relates to. 
* gganimate solves this by providing a set of variables for each frame, which can be inserted into plot labels using glue syntax.
* Use glue syntax to insert frame variables in plot labels and titles.
* Different transitions provide different frame variables. closest_state only makes sense for transition_states() and is thus only available when that transition is used.
```{r}
anim + 
  ggtitle('Now showing {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
```
## Line plot equivalent of above animate
```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_line(aes(group = rep(1:50, 3)), colour = 'grey') + 
  geom_point()
```
## Add some aesthetics that distinguish the different species

```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species)) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1) +
   ggtitle('Now showing {closest_state}',
          subtitle = 'Frame {frame} of {nframes}')
```
## Set the group directly
* The group aesthetic defines how the data in a layer is matched across the animation.
* 
```{r}

seq_along(iris$Species)

ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(group = seq_along(Species))) + 
  transition_states(Species,
                    transition_length = 1,
                    state_length = 0.5)
```
## Group enhancement
```{r}
ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species, group = 1L)) + 
  transition_states(Species,
                    transition_length = 2,
                    state_length = 1)
```
## Enter and Exit
* modify the state of appearing (entering) and disappearing (exiting) data, so that the animation can tween from and to the new state.
```{r}
anim <- ggplot(iris, aes(x = Petal.Width, y = Petal.Length)) + 
  geom_point(aes(colour = Species), size = 2) + 
  transition_states(Species,
                    transition_length = 1,
                    state_length = 0.5)

anim + 
  enter_fade() + 
  exit_shrink()
```
## enter and Exit through drift
```{r}
anim + 
  enter_fade() + enter_drift(x_mod = -1) + 
  exit_shrink() + exit_drift(x_mod = 5)
```



## Additional readings
* https://exts.ggplot2.tidyverse.org/gallery/
* https://www.r-graph-gallery.com

## Reference:
* https://cran.r-project.org/web/packages/gganimate/vignettes/gganimate.html

