---
title: "Visualisation (ggplot)"
author: "Revendranath T"
date: "27/9/2021"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Standard plots

```{r cars, echo = TRUE}
summary(cars)
plot(cars)
```

## Prerequisites
```{r}
library(tidyverse)
#install.packages("tidyverse")
#library(tidyverse)
```

## About mpg dataset
* https://archive.ics.uci.edu/ml/datasets/Auto%2BMPG
* https://www.kaggle.com/uciml/autompg-dataset

```{r}
# Look at this dataframe
mpg %>% glimpse()
str(mpg)
```

```{r}
fct_unique(as_factor(mpg$class))

fct_unique(as_factor(mpg$drv))
```

### First plot using ggplot
```{r}
#What does this do? What do you think of the code?
mpg %>% ggplot()  #--> ggplot2

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

p1 <- ggplot(data = mpg) 

p1

p2 <-  geom_point(mapping = aes(x = displ, y = hwy))
p2

p1
p2

p1 + p2
```

## Standard graph template
```{r}

ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

### Aesthetics mapping
* color & size
```{r}
#class
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, colour = class))
```

```{r}
#size
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))
```

```{r}
# alpha
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

```{r}
# shape
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

```{r}
#all plots in blue color
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```


```{r}
#change to red
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class), color = "red")
```


```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, 
                           y = hwy, 
                           alpha = class, 
                           color = class))
```

```{r}

mpg %>% summary()

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, 
                           y = hwy, 
                           shape = as_factor(cyl), 
                           color = as_factor(cyl)))

```


## Facets
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ as_factor(cyl), nrow = 2)
```
```{r}
mpg %>% glimpse()

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = cty)) + 
  facet_wrap(~ as_factor(cyl), nrow = 2)
```


```{r}

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 3)
```

### Geometric objects
```{r}
# points
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```
```{r}
ggplot(data = mpg) + 
 geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
#smoothing
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))  # scatter plot using geom_point function
+  geom_smooth(mapping = aes(x = displ, y = hwy))

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))

#loess
```

```{r}
mpg %>% glimpse()

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))

#smoothing 
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, 
                           y = hwy, 
                           shape = as_factor(cyl), 
                           color = as_factor(cyl))) +
  geom_smooth(mapping = aes(x = displ, y = hwy, linetype = drv))

+
  facet_wrap(~, nrow = 4 )
#separates the cars into three lines based on their drv value, which describes a car’s drivetrain. One line describes all of the points with a 4 value, one line describes all of the points with an f value, and one line describes all of the points with an r value. Here, 4 stands for four-wheel drive, f for front-wheel drive, and r for rear-wheel drive.
```

#Check the differences in these smoothing
```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
              
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
    
ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv),
    show.legend = FALSE
  )

ggplot(data = mpg) +
  geom_smooth(
    mapping = aes(x = displ, y = hwy, color = drv)
  )
```

```{r}
#multiple geoms on same plot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = drv, color = drv)) +
  geom_smooth(mapping = aes(x = displ, y = hwy, color = drv))
```

```{r}
#same as above?
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, shape = drv, color = drv)) + 
  geom_point() + 
  geom_smooth()
```


```{r}
#How is this different?
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth()
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class),
    show.legend = FALSE) + 
  geom_smooth(mapping = aes(color = drv),
    show.legend = FALSE)

#Think how you can visualise your plot
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "subcompact"), se = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point(mapping = aes(color = class)) + 
  geom_smooth(data = filter(mpg, class == "suv")) #, se = FALSE)
```
```{r}
diamonds %>% glimpse()

fct_count(diamonds$cut)
```

#Statistical transformations
```{r}
#Bar chart
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
#Better way to do it
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
```{r}
#another example
demo <- tribble(
  ~cut,         ~freq,
  "Fair",       1610,
  "Good",       4906,
  "Very Good",  12082,
  "Premium",    13791,
  "Ideal",      21551
)

#frequency
ggplot(data = demo) +
  geom_bar(mapping = aes(x = cut, y = freq), stat = "identity")
```

```{r}
#proportions
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = stat(prop), group = 1))
```

#summaries with stat_summary()
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
    #, fun = mean
  )

```
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = price),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
diamonds %>% glimpse()
levels(diamonds$clarity)

diamonds$clarity %>% head()
```

### Position adjustments
```{r}
diamonds %>% glimpse()
#Check the difference between p1 & p2
p1 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))

p1
```

```{r}

p2 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
p2
```

```{r}
#Can we do better?
diamonds %>% select(clarity) %>% unique()

p3 <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
p3
```

```{r}
fct_count(diamonds$cut)
```


```{r}
diamonds %>% select(clarity, cut) %>% 
  group_by(cut, clarity) %>% 
  summarise(count_x = n())

fct_count(diamonds$clarity)
```


```{r}
#position = "fill" works like stacking, but makes each set of stacked bars the same height. This makes it easier to compare proportions across groups.
#stacked plots
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```

```{r}
#position identity
ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(position = "identity")

# or
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "identity")
```

```{r}

ggplot(data = diamonds, mapping = aes(x = cut, fill = clarity)) + 
  geom_bar(alpha = 1/5, position = "identity")
```

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, colour = clarity)) + 
  geom_bar(fill = NA, position = "identity")
```

```{r}
#position = "dodge" places overlapping objects directly beside one another. This makes it easier to compare individual values.

ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge") 
```

### Coordinate systems
* coord_flip() switches the x and y axes. This is useful (for example), if you want horizontal boxplots. It’s also useful for long labels: it’s hard to get them to fit without overlapping on the x-axis.

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge") +
  coord_flip()
```

```{r}
data(mpg)
```


```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot()
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()

ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = class, y = hwy)) +
  coord_flip()
```

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + 
  geom_boxplot() +
  coord_flip()
```

* coord_quickmap() sets the aspect ratio correctly for maps. This is very important if you’re plotting spatial data with ggplot2 (which unfortunately we don’t have the space to cover in this book).

```{r}
library(maps)
nz <- map_data("nz")
nz %>%  glimpse()
```

```{r}
ggplot(nz, aes(x=long, y=lat, group = group)) +
  geom_polygon(fill = "white", colour = "black")
```

```{r}
ggplot(nz, aes(x=long, y=lat, group = group)) +
  geom_polygon(fill = "white", colour = "black") +
  coord_quickmap()
```


* coord_polar() uses polar coordinates. Polar coordinates reveal an interesting connection between a bar chart and a Coxcomb chart.
```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar
bar + coord_flip()
bar + coord_polar()
```

## Layers of grammer of graphics
```{r}
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION1>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
    <GEOM_FUNCTION2>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

## Esquisse
```{r}
library(esquisse)
```

```{r}
getwd()
```


```{r}
#ggplot already loaded
#library(ggplot2)
library(maptools)
library(rgeos)
library(ggmap)

library(scales)
library(RColorBrewer)
set.seed(8000)

# A shapefile requires the .shp as well as the .shx and .dbf files.
# Download all of them
#shp <- readShapeSpatial('C:/Users/user/Google Drive/RWorkshop/India_State_Boundary.shp')

shp <- readShapeSpatial('G:/My Drive/Courses/Programming for Analytics (MPBA G507)/For Students/Course materials/Visualization/India_State_Boundary.shp')

shp <- readShapeSpatial('G:/My Drive/Courses/Programming for Analytics (MPBA G507)/For Students/Course materials/6) Visualization')

class(shp)
shp %>% glimpse()
plot(shp)
names(shp)
print(shp$State_Name)
```

```{r}
#library(datapasta)

ind_unemp <- data.frame(
  stringsAsFactors = FALSE,
       check.names = FALSE,
  States = c("Andaman & Nicobar", 
             "Andhra Pradesh",
             "Arunachal Pradesh",
             "Assam",
             "Bihar",
             "Chandigarh",
             "Chhattishgarh",
             "Daman and Diu and Dadra and Nagar Haveli",
             "Delhi",
             "Goa",
             "Gujarat",
             "Haryana",
             "Himachal Pradesh",
             "Jammu and Kashmir",
             "Jharkhand",
             "Karnataka",
             "Kerala",
             "Ladakh",
             "Lakshadweep",
             "Madhya Pradesh",
             "Maharashtra",
             "Manipur",
             "Meghalaya",
             "Mizoram",
             "Nagaland",
             "Odisha",
             "Puducherry",
             "Punjab",
             "Rajasthan",
             "Sikkim",
             "Tamilnadu",
             "Telengana",
             "Tripura",
             "Uttar Pradesh",
             "Uttarakhand",
             "West Bengal"
             ),
          Unemployment = c(8.7, 
                           1.9,
                           13,
                           3.2,
                           10.7,
                           21.4,
                           1.8,
                           28.1,
                           15,
                           15.4,
                           9.4,
                           2,
                           7.5, 
                           15.4,
                           2.2,
                           4.3,
                           0.4,
                           1.1,
                           9.6,
                           6.8,
                           21,
                           0,
                           4.8,
                           4,
                           13.2,
                           5,
                           3.1,
                           7.4,
                           9.6,
                           6.8,
                           21,
                           0,
                           4.8,
                           4,
                           13.2,
                           5)
)


class(shp)
shp.f <- fortify(shp, region = "State_Name")
class(shp.f)

shp.f %>% glimpse()

shp.f %>% select(id) %>%  unique()

ind_unemp %>% glimpse()

summary(ind_unemp)

typeof(ind_unemp)

merge.shp.coef <- shp.f %>%  full_join(ind_unemp, by =c("id" = "States"))

merge.shp.coef %>% glimpse()

#merge.shp.coef$order
final.plot <- merge.shp.coef[order(merge.shp.coef$order), ]

final.plot %>% glimpse()
```

```{r}

# NZ
# ggplot(nz, aes(long, lat, group = group)) +
#   geom_polygon(fill = "white", colour = "black") +
#   coord_quickmap()
                   
#India
ggplot(data = final.plot) + 
  geom_polygon(aes(x = long, 
                   y = lat, 
                   group = group, 
                   fill = Unemployment), 
               color = "black", 
               size = 0.25) + coord_map()

#ggsave("India_IMR_2013_BLUE.png",dpi = 300, width = 20, height = 20, units = "cm")

ggplot() +
  geom_polygon(data = final.plot,
               aes(x = long, y = lat, group = group, fill = count), 
               color = "black", size = 0.25) + 
  coord_map()+
  scale_fill_distiller(name="IMR", palette = "YlGn")+
  labs(title="IMR of Different States of India 2013")+
  xlab('Longitude')+
  ylab('Latitude')
ggsave("India_IMR_2013_YlGn.png",dpi = 300, width = 20, height = 20, units = "cm")

ggplot() +
  geom_polygon(data = final.plot,
               aes(x = long, y = lat, group = group, fill = count), 
               color = "black", size = 0.25) + 
  coord_map()+
  scale_fill_gradient(name="IMR", limits=c(0,100), low = 'white', high = 'red')+
  labs(title="IMR of Different States of India 2013")+
  xlab('Longitude')+
  ylab('Latitude')
ggsave("India_IMR_2013_NO_ST_NAME.png",dpi = 300, width = 20, height = 20, units = "cm")

cnames <- aggregate(cbind(long, lat) ~ id, data=final.plot, FUN=function(x) mean(range(x)))

ggplot() +
  geom_polygon(data = final.plot,
               aes(x = long, y = lat, group = group, fill = count), 
               color = "black", size = 0.25) + 
  coord_map()+
  scale_fill_gradient(name="IMR", limits=c(0,100), low = 'white', high = 'red')+
  labs(title="IMR of Different States of India 2013")+
  xlab('Longitude')+
  ylab('Latitude')+
  geom_text(data=cnames, aes(long, lat, label = id), size=3, fontface="bold")
ggsave("India_IMR_2013.png",dpi = 300, width = 20, height = 20, units = "cm")

```


