---
title: "Introduction & Basics of Strings"
author: "PfA Teaching Team"
date: "9/30/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction & Basics of Strings

## Attach tidyverse meta-package
```{r}
library(tidyverse)
```

```{r}
s1 <- "Welcome to BITS Pilani!"
nchar(s1)
s2 <- 'Insert "double quote" between two single quotes'
s3 <- "Insert \"double quote\" between two single quotes"
typeof(s1)
print(s1)
print(s2)
s2
s3

s4 <- "Sagar Kapoor"
nchar(s4)

s5 <- "Sagar \"MBA\" Kapoor"
nchar(s5)

#or
s6 <- 'Sagar "MBA" Kapoor'
s6
nchar(s6)
```
```{r}
double1 <- "\"" 
# or 
double2 <- '"'
print(double1)
print(double2)
```

```{r}

single1 <- '\'' 
# or 
single2 <- "'"

print(single1)
print(single2)
```


## What happens when you assign an incomplete string?
* Try this on console
```{r}
s3 <- "Incomplete quote example
```

## Special characters
* Handful of special characters exist
```{r}
print("\u100")

print("\u1138")
print("\u4144")
print("\u9999")
```
```{r}
print("\\u100")
s1 <-  "\\u100"
print(s1)
print(nchar(s1))
```


## Combine strings using c() and str_c()
```{r}
# Combine multiple strings into a vector
c("Hello", "Welcome", "to", "BITS", "Pilani")

# Combine multiple strings into one string using str_c
str_c("Hello,", "Welcome", "to", "BITS", "Pilani!")

# Combine multiple strings into one string using str_c with a separator
str_c("Hello,", "Welcome", "to", "BITS", "Pilani!", sep = ",")

str_c("Hello,", "Welcome", "to", "BITS", "Pilani!", sep = " ")
```
## Handling NAs (or missing values): str_replace_na()
```{r}
greeting1 <- c("Welcome to BITS Pilani")
name1 <- "Sagar"
str_c("Hello ", name1, "! ", greeting1, "!")

name2 <- NA
str_c("Hello ", name2, "! ", greeting1, "!")

#if you find missing value, handle the same using: str_replace_na()
str_c("Hello ", str_replace_na(name2), "! ", greeting1, "!")
str_c("Hello ", str_replace_na(name2, ""), "! ", greeting1, "!")
str_c("Hello ", str_replace_na(name2, "There"), "! ", greeting1, "!")
nchar(name2)
length(name2)
```
## Vectorised operations happen within str_c()
```{r}
# Illustration 1
str_c("before-", c("one", "two", "three"), "-after")

names <- c("Sagar", "Sougandh", "Ajay", NA, "Indika", "Francis", "Pratik")
str_c("Hello ", str_replace_na(names, ""), "! ", greeting1, "!")
```

```{r}

# Illustration 2
greeting <- "Welcome to "
course <- "Programming for Analytics!"
thursday <- FALSE
tuesday <- TRUE


str_c(
  greeting, " ", course, 
  if (thursday) " Lab Session on Thursday",
  ".")

str_c(
  greeting, " ", course, 
  if (tuesday) " Lab Session on Tuesday",
  ".")

friday <- TRUE

str_c(
  greeting, " ", course, 
  if (friday) " Lecture Session on Friday",
  ".")

```
## Check how collapse is useful
```{r}
clean_vec <- c("Welcome to ", "BITS Pilani", "!")
print(length(clean_vec))
print(class(clean_vec))
print(paste0(clean_vec))
```

```{r}
str_c_vec <- str_c(c("Welcome to ", "BITS Pilani", "!"))
print(length(str_c_vec))
print(class(str_c_vec))
print(paste0(str_c_vec))
```
## use collapse
```{r}
c_collapse <-  str_c(c("Welcome to ", "BITS Pilani", "!"), collapse = ", ")
print(length(c_collapse))
print(class(c_collapse))
print(paste0(c_collapse))
```

```{r}
collapse <- str_c("Welcome to ", "BITS Pilani", "!", collapse = ", ")
print(length(collapse))
print(class(collapse))
print(paste0(collapse))
```

## Get string length using str_length()
```{r}
#length of a string
str_length("Welcome to BITS Pilani!")

#length of a  multiple strings
str_length(c( "Hello!", "Welcome to BITS Pilani!"))
```

## Sub-setting Strings
* Extract a portion of string starting from position "m" & ending at postion "n"
```{r}
x <- c("BITS", "Pilani", "Rajasthan")
str_sub(x, 1, 3)
```

* Extract last position to backward values
```{r}
str_sub(x, -3, -1)



```

## print the reverse of strings in a character
```{r}
str_sub(x[1], -4, -1)

str_sub(x, -1, -str_length(x))
arg1 <- c(-1, -str_length(x[1]))
arg2 <- c(-1, -str_length(x[2]))
arg3 <- c(-1, -str_length(x[3]))

x %>% invoke_map(c(arg1,arg2,arg3), str_sub) #more needs to be done on logic
print(x)
print(stringi::stri_reverse(x)) #this works, but logic codes needs to be written
```

## str_sub() does not throw an error when string length is small
```{r}
str_sub("BITS", 1, 100)
```
## Convert strings to lower/upper cases

```{r}
print(x)
print(str_sub(x, 1, 1))
```

```{r}
# Upper to lower (First character)
str_sub(x, 1, 1) <- str_to_lower(str_sub(x, 1, 1))
x

# Lower to Upper (Third character)
str_sub(x, 3, 3) <- str_to_upper(str_sub(x, 3, 3))
x
```
## Be aware of the local language you are operating your strings
```{r}
# default assumes English locale
x <- c("apple", "eggplant", "banana")
str_sort(x)
```
```{r}
#English locale
str_sort(x, locale = "en")  
```

```{r}
# in Hawaiian, e comes before b
str_sort(x, locale = "haw") # Hawaiian
```

