---
title: "Untitled"
author: "PfA Teaching Team"
date: "9/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Key operations on text data through pattern matching

* Determine which strings match a pattern.
* Find the positions of matches.
* Extract the content of matches.
* Replace matches with new values.
* Split a string based on a match.

```{r}
library(tidyverse)
```
## Detect Matches
* To verify if a character vector matches a pattern
* use str_detect()
* returns a logical vector the same length as the input
```{r}
x <- c("BITS Pilani", "MBA BA", " Winter is coming")
x
#Is there "e" in vector x?
str_detect(x, "e")

#How about "i" in vector x?
str_detect(x, "i")
#or
str_detect(x, regex("i"))

#ignore case sensitivity
str_detect(x, regex("i", ignore_case = TRUE))

str_detect(x, regex("A", ignore_case = TRUE))

```
## How many words in "words" vector starts with s, t, h, and ends with a,e,i,o,u?

```{r}
words
```

```{r}
# Note:  logical vector in a numeric context, FALSE becomes 0 and TRUE becomes 1
sum(str_detect(words, regex("^s", ignore_case = TRUE)))
#or
sum(str_count(words, regex("^s", ignore_case = TRUE)))

sum(str_detect(words, regex("^t", ignore_case = TRUE)))
#or
sum(str_count(words, regex("^t", ignore_case = TRUE)))

sum(str_detect(words, regex("^h", ignore_case = TRUE)))
#or
sum(str_count(words, regex("^h", ignore_case = TRUE)))
```

```{r}

#ending with h
sum(str_count(words, regex("h$", ignore_case = TRUE)))

#starting with y and ending with y: yesterday
sum(str_count(words, regex("^y.*.y$", ignore_case = TRUE)))

words[str_detect(words, regex("^y.*.y$", ignore_case = TRUE))]


sum(str_count(words, regex("^w.*.w$", ignore_case = TRUE)))
words[str_detect(words, regex("^w.*.w$", ignore_case = TRUE))]


sum(str_count(words, regex("^s.*.t$", ignore_case = TRUE)))
words[str_detect(words, regex("^a.*.y$", ignore_case = TRUE))]

```


```{r}
# Together

sum(str_detect(words, regex("^[s,t,h]", ignore_case = TRUE))) #^[s,t,h] implies first character starting with s or t or h
#or
sum(str_count(words, regex("^[s,t,h]", ignore_case = TRUE)))

# ends with a,e,i,o,u
sum(str_detect(words, regex("[a,e,i,o,u]$", ignore_case = TRUE)))
#or
sum(str_count(words, regex("^[a,e,i,o,u]$", ignore_case = TRUE)))
# instead of sum, you may perform mean
mean(str_detect(words, regex("[a,e,i,o,u]$", ignore_case = TRUE))) 
#271/980

test_words <- c("y", "yesterday")
str_detect(test_words, regex("^y$", ignore_case = TRUE))
```

## find patterns starting with "qui"
```{r}
sum(str_count(words, regex("^qui", ignore_case = TRUE)))

words[str_detect(words, regex("^qui", ignore_case = TRUE))]
```
```{r}
words[str_detect(words, regex("^qui", ignore_case = TRUE))]
```

```{r}
# excludes all the words containing with q, or u, or i, or a
words[str_detect(words, regex("^[^quia]+", ignore_case = TRUE))]
```

```{r}
x <- c("BITS", "Pilani", "MBA", "XPQ", "TTT") 
str_detect(x, regex("[aeiou]", ignore_case = TRUE))

x[str_detect(x, regex("[aeiou]", ignore_case = TRUE))]

str_detect(x, regex("[a,e,i,o,u]", ignore_case = TRUE))

str_detect(x, regex("^[^aeiou]+$", ignore_case = TRUE))

str_detect(x, regex("^[^aeiou]+", ignore_case = TRUE))

```


```{r}
# Find all words containing at least one vowel, and negate
words[str_detect(words, "[aeiou]")]

!str_detect(words, "[aeiou]")
words[!str_detect(words, "[aeiou]")]

no_vowels_1 <- !str_detect(words, "[aeiou]")

# Find all words consisting only of consonants (non-vowels)
str_detect(words, "^[^aeiou]+$")

words[str_detect(words, "^[^aeiou]+$")]

no_vowels_2 <- str_detect(words, "^[^aeiou]+$")

#Validation of two approaches
identical(no_vowels_1, no_vowels_2)

```

## Exact Matches
* Find "to", "be", "the", and "a" in following quotes
```{r}
quote1 <- "Strive the not to be a success, but rather to be of value begin together"
quote1_author <- "Albert Einstein"
quote2 <- "Twenty years from now you will be more disappointed by the things that you didn’t do than by the ones you did do"
quote2_author <- "Mark Twain"

articles <- c("a", "the")
prepos_verbs <- c("to", "be")

articles_match <- str_c(articles, collapse = "|" )
prepos_verbs_match <- str_c(prepos_verbs, collapse = "|")
```

```{r}
# str_extract() gets only the first match
quote1_has_articles <- str_extract(quote1, articles_match) 
# or 
quote1_has_articles <- str_extract(quote1, "a|the")

a_vec = c("a", "the", "an")

str_extract(quote1, "a OR the")

str_extract_all(quote1, a_vec)

quote1_has_prepos_verbs <- str_extract(quote1, prepos_verbs_match)
```

```{r}

quote2_has_articles <- str_extract(quote2, articles_match)
quote2_has_prepos_verbs <- str_extract(quote2, prepos_verbs_match)

# str_extract_all() to get all the matches
quote1_has_articles <- str_extract_all(quote1, articles_match)
quote1_has_prepos_verbs <- str_extract_all(quote1, prepos_verbs_match)


quote2_has_articles <- str_extract_all(quote2, articles_match)
quote2_has_prepos_verbs <- str_extract_all(quote2, prepos_verbs_match)

# Pass argument simplify = TRUE to return matrix with short matches
str_extract_all(quote2, articles_match, simplify = TRUE)
str_extract_all(quote2, prepos_verbs_match, simplify = TRUE)
```

## Grouped Matches
* Assume that you need to identify nouns in a sentence. What's your approach?
* Hint: Nouns come after the articles (a, the)
```{r}
noun <- "(a|the) ([^ ]+)"

quote1_has_noun <- str_extract_all(quote1, noun)
#or
quote1_has_noun <- quote1 %>% str_extract_all(noun)


quote2_has_noun <- str_extract_all(quote2, noun)
#or
quote2_has_noun <- quote2 %>% str_extract_all(noun)

```
## str_match(): provides individual component from the match in the form of a matrix with one column for the complete match followed by one column for each group
```{r}
quote1_has_noun %>% str_match(noun)

quote1_has_noun %>% str_match_all(noun)

quote2_has_noun %>% str_match(noun)

quote2_has_noun %>% str_match_all(noun)

 
c(quote1_has_noun, quote2_has_noun) %>% str_match(noun)

c(quote1_has_noun, quote2_has_noun) %>% str_match_all(noun)
```

## Splitting
* str_split() : splits a string up into pieces
```{r}
x <- c("AA", "C", "BB", "CC", "DD", "CCCCC", "CCC", "ABBCCCDDDEEEE")

fruit <- c("apple", "banana", "pear", "pineapple", "frooty")

x %>% str_split(" ")

fruit %>% str_split(" ")
```

```{r}
"a|b|c|d" %>% 
  str_split("\\|") %>% 
  .[[1]]
```
## set maximum number of splits
```{r}
x <- c("Program: MBA: ABC", "Specialisation: BA: ABC", "University: BITS: ABC", "Place: Pilani: ABC")
x %>% str_split(": ", simplify = TRUE)
x %>% str_split(": ", n = 2, simplify = TRUE)
```
## Split the sentences by words
```{r}
print(quote1)
print(quote2)
```

```{r}
quote1 %>%  str_view_all(boundary("word"))
quote1 %>%  str_split(boundary("word"))
```

```{r}
quote2 %>%  str_view_all(boundary("word"))
quote2 %>%  str_split(boundary("word"))

```
## Find exact matches and their positions
* str_locate()
* str_locate_all()
```{r}
x <- "BITS Pilani Rajasthan"
length(x)
str_length(x)
```

```{r}
# find the positions of "i"
str_locate(x, "i")
str_locate_all(x, "i")

str_locate_all(x, regex("i", ignore_case = TRUE))
```

```{r}

# find the positions of "a"
str_locate(x, "a")
str_locate_all(x, "a")
```
## Further readings
* Check https://stringr.tidyverse.org/reference/index.html
* Check stringi package
