---
title: "Regex - Pattern Matching"
author: "PfA Teaching Team"
date: "10/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pattern Matching using Regular Expressions (Regex)
* Enables identificatoin of patterns in strings
* Key functions used at the beginning are:
  * str_view() and
  * str_view_all()
  * Both functions take character vector and regular expressions, and describe the patterns within

## Basic Illustrations
```{r}
x <- c("BITS Pilani", "BITS Hyderabad", "BITS Goa")
# Is there any matching here?
str_view(x, "bits")
```

```{r}
x <- c("BITS Pilani", "BITS Hyderabad", "BITS Goa")
#how about matching here?
str_view(x, "BITS")
```
## Find "a"
```{r}
str_view(x, "a" )
```

## Find all "a"s occuring more than once
```{r}
str_view_all(x, "a")
```


## Match any character before and after, except newline: use . (i.e., dot)
```{r}
str_view(x, ".I.")
```
## Anything before "a"
```{r}
str_view(x, ".a")
```

## How to find . in text?
* Use escape through double backslashes \\
```{r}
xdot <- "Hello.Welcome.To.BITS.Pilani!"
str_view(xdot, "\\.")
```

## Find positions of all dots
```{r}
str_view_all(xdot, "\\.")
```


## What if you need to find "\" in text?
* use \\\\
```{r}
x <- "BITS\\Pilani" # creates a string BITS\Pilani
print(x)
x
str_view(x, "\\\\")
```

## Anchors
*
```{r}
quote1 <- "Strive not to be a success, but rather to be of value"
quote1_author <- "Albert Einstein"
quote2 <- "Twenty years from now you will be more disappointed by the things that you didn’t do than by the ones you did do"
quote2_author <- "Mark Twain"

```
## Identify occurances of "e" and "s" in quote1
```{r}
#occurances of e
str_view_all(quote1, "e")
#occurances of S
str_view_all(quote1, "s")
```
## What if I want to see if a sentence is 
* ending with "e" or 
* beginning with "S"

## It requires anchors

  * ^ to match the start of the string.
  * $ to match the end of the string.

```{r}
str_view(quote1, "e$")
```

```{r}
str_view(quote1, "^S")
```

## Four Strings, and you are searching for a string with exact match.
* use both ^ & $
```{r}
x <- c("BITS Pilani", "BITS Hyderabad", "BITS", "BITS Goa")

str_view(x,"BITS")
str_view(x,"^BITS$")
#or
str_view_all(x,"BITS")
str_view_all(x,"^BITS$")
```
## Matching more than one character
## Four ways
* \d: matches any digit.
* \s: matches any whitespace (e.g. space, tab, newline).
* [abc]: matches a, b, or c.
* [^abc]: matches anything except a, b, or c.

# Find "n" between "a" and "d"
```{r}
# does this work?
str_view(c("abd", "and", "a*d", "a.d"), "a[.]d")

# ofcourse
str_view(c("abd", "and", "a*d", "a.d"), "and")
str_view(c("abd", "and", "a*d", "a.d"), "a[n]d")
```
## What if "abd" and "and" to be selected below?
```{r}
str_view(c("abd", "and", "a*d", "a.d"), "a[b,n]d")
# or
#pipe operator within paranthesis also works
str_view(c("abd", "and", "a*d", "a.d"), "a(b|n)d")
```


### Suppose you want to find * and . between a and d
```{r}

str_view(c("abd", "and", "a*d", "a.d"), "a[*,.]d")

str_view(c("abd", "and", "a*d", "a.d"), "a[\*,\.]d") #Single escape character doesn't work

#use the following than previous one without escape character
str_view(c("abd", "and", "a*d", "a.d"), "a[\\*,\\.]d")
```
## Repetitive pattern matching
use the following symbols for
* ?: 0 or 1 repititions
* +: 1 or more repititions
* *: 0 or more repititions

```{r}
x <- c("AA", "C", "BB", "CC", "DD", "CCCCC", "CCC", "ABBCCCDDDEEEE")
str_view(x, "CC?")
str_view(x, "CC+")
str_view(x, "CC*")

str_view(x, 'CC[D]+')
```

### Specify the number of matches precisely
* {n}: exactly n
* {n,}: n or more
* {,m}: at most m
* {n,m}: between n and m
```{r}
x <- c("AA", "C", "BB", "CC", "DD", "CCCCC", "CCC", "ABBCCCDDDEEEE")
# Exactly n number of matches
str_view(x, "C{1}")
str_view(x, "C{2}")
str_view(x, "C{3}")
str_view(x, "C{5}")
```

# n, or more than n number of matches
```{r}
# 2, or more than 2 number of matches
str_view(x, "C{2,}")

# 3, or more than 3 number of matches
str_view(x, "C{3,}")
```

# utmost "n" number of matches
```{r}
# utmost "2" number of matches
# implies min:1 & max:2 matches
str_view(x,"C{1,2}")

# utmost "3" number of matches
# implies min:1 & max:3 matches
str_view(x,"C{1,3}")
```

## str_subset() & str_which()
```{r}
fruit <- c("apple", "banana", "pear", "pineapple", "frooty")

str_subset(fruit, "a")

str_which(fruit, "a")

str_subset(fruit, "^b")

# Returns elements that do NOT match
str_subset(fruit, "^p", negate = TRUE)

# Missings never match
str_subset(c("a", NA, "b"), ".")

```


