---
title: "Shiny Examples"
author: "Revendranath T"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Free text

```{r}
library(shiny)

ui <- fluidPage(
  textInput("name", "What's your name?"),
  passwordInput("password", "What's your password?"),
  textAreaInput("story", "Tell me about yourself", rows = 3)
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)

```

## Numeric inputs
```{r}
library(shiny)

ui <- fluidPage(
  numericInput("num", "Number one", value = 25, min = 0, max = 100),
  numericInput("num1", "Number one.1", value = 50,  min = 0, max = 100),
  sliderInput("num2", "Number two", value = 40, min = 0, max = 100),
  sliderInput("stu_rng", "Student Range", value = c(17, 20), min = 17, max = 45),
  sliderInput("emp_rng", "Employee Range", value = c(22, 30), min = 18, max = 65),
  sliderInput("ret_rng", "Retirement Range", value=60, min = 60, max = 100)
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## Dates as Inputs
```{r}
library(shiny)

ui <- fluidPage(
  dateInput("dob", "When were you born?"),
  dateRangeInput("holiday", "When do you want to go on vacation next?")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Restrict the choices

```{r}
library(shiny)

animals <- c("dog", "cat", "mouse", "bird", "other", "I hate animals")

ui <- fluidPage(
  selectInput("state", "What's your favourite state?", state.name),
  radioButtons("animal", "What's your favourite animal?", animals),
  selectInput("state", "What's your second favourite animal?", animals)
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Radio buttons with smileys
```{r}
library(shiny)

ui <- fluidPage(
  radioButtons("rb", "Choose one:",
    choiceNames = list(
      icon("angry"),
      icon("smile"),
      icon("sad-tear")
    ),
    choiceValues = list("angry", "happy", "sad")
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Multiple dropdown Inputs for selection
```{r}
library(shiny)

ui <- fluidPage(
  selectInput(
    "state", "What's your favourite state?", state.name,
    multiple = TRUE
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## Checkboxes or multiple values for radio buttons
```{r}
library(shiny)

ui <- fluidPage(
  checkboxGroupInput("animal", "What animals do you like?", animals),
  checkboxGroupInput("animal1", "2nd fav animal?", animals)
  
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

# Select one checkbox by default
```{r}
library(shiny)

ui <- fluidPage(
  checkboxInput("cleanup", "Clean up?", value = TRUE),
  checkboxInput("shutdown", "Shutdown?"),
  checkboxInput("good", "Session good?", value = TRUE),
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Upload files
* use fileInput
```{r}
ui <- fluidPage(
  fileInput("upload", NULL),
  actionButton("submit", "Submit File")
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Action Buttons
* actionButton() or 
* actionLink() : Actions links and buttons are most naturally paired with observeEvent() or eventReactive() in  server function.

```{r}
library(shiny)

ui <- fluidPage(
  actionButton("click", "Click me!"),
  actionButton("drink", "Drink me!", icon = icon("cocktail"))
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

# Customise the button
* customise the appearance using the **class** argument by using one of "btn-primary", "btn-success", "btn-info", "btn-warning", or "btn-danger". 
* change the size with "btn-lg", "btn-sm", "btn-xs". 
* make buttons span the entire width of the element they are embedded within using "btn-block"

```{r}


ui <- fluidPage(
#   fluidRow(
#     actionButton("click", "Click me!", class = "btn-danger"),
#     actionButton("drink", "Drink me!", class = "btn-lg btn-success")
#   ),
#   fluidRow(
#     actionButton("eat", "Eat me!", class = "btn-block")
#   )
# )
  fluidRow(
  column(4,
    actionButton("click", "Click me!", class = "btn-danger"),
    actionButton("drink", "Drink me!", class = "btn-lg btn-success")
  ),
  column(8,
    actionButton("click", "Click me!", class = "btn-danger"),
    actionButton("drink", "Drink me!", class = "btn-lg btn-success")
  )
  ),
  fluidRow(
  column(6,
    actionButton("eat", "Eat me!", class = "btn-danger")
  ),
  column(6,
    actionButton("eat", "Eat me!", class = "btn-danger")
  )
  )
)

  
server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## Outputs
* Each output function on the front end is coupled with a render function in the back end
* Three types
  * text, 
  * tables, and 
  * plots.
  
## Text
* Output regular text with textOutput() and fixed code and console output with verbatimTextOutput().
```{r}
library(shiny)

ui <- fluidPage(
  textOutput("text"),
  verbatimTextOutput("code")
  #textOutput("code")
)

server <- function(input, output, session) {
  output$text <- renderText({ 
    "Welcome to Shiny!" 
  })
  output$code <- renderPrint({ 
    summary(1:10) 
  })
}

shinyApp(ui, server)

# print( "Welcome to Shiny!")
# summary(1:10)
```

* Note that the {} are only required in render functions if need to run multiple lines of code. 
* renderText() combines the result into a single string, and is usually paired with textOutput()
* renderPrint() prints the result, and is usually paired with verbatimTextOutput()

```{r}

ui <- fluidPage(
  textOutput("text"),
  verbatimTextOutput("code")
)

server <- function(input, output, session) {
  output$text <- renderText("Welcome to Shiny!")
  output$code <- renderPrint(summary(1:10))
}

shinyApp(ui, server)
```

## Tables
* tableOutput() and renderTable() render a static table of data, showing all the data at once.
* tableOutput() is most useful for small, fixed summaries (e.g. model coefficients); * dataTableOutput() and renderDataTable() render a dynamic table, showing a fixed number of rows along with controls to change which rows are visible.
* dataTableOutput() is most appropriate if you want to expose a complete data frame to the user

```{r}
ui <- fluidPage(
  tableOutput("static"),
  dataTableOutput("dynamic")
)

server <- function(input, output, session) {
  output$static <- renderTable(head(mtcars))
  output$dynamic <- renderDataTable(mtcars, options = list(pageLength = 11))
}

shinyApp(ui, server)
```


## Plots
* Any type of R graphic (base, ggplot2, or otherwise) with plotOutput() and renderPlot()
```{r}
ui <- fluidPage(
  plotOutput("plot", width = "400px")
)
server <- function(input, output, session) {
  output$plot <- renderPlot(plot(1:5), res = 96)
}

shinyApp(ui, server)
```


## Notes:
* The render function does two things:
  * It sets up a special reactive context that automatically tracks what inputs the output uses.
  * It converts the output of your R code into HTML suitable for display on a web page.

##  Reactive programming
* The big idea in Shiny: you don’t need to tell an output when to update, because Shiny automatically figures it out for you.
```{r}
ui <- fluidPage(
  textInput("name", "What's your name?"),
  textInput("fav_song", "What's your favourite song?"),
  textOutput("greeting"),
  textOutput("song")
)

server <- function(input, output, session) {
  output$greeting <- renderText({
    paste0("Hello ", input$name, "!")
  })
  output$song <- renderText({
    paste0("Your favourite song is ", input$fav_song, "!")
  })
}

shinyApp(ui, server)
```


