---
title: "Page Layouts"
author: "Revendranath T"
date: "8/20/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(shiny)

ui <- fluidPage(
  titlePanel("Central limit theorem"),
  sidebarLayout(
    sidebarPanel(
      numericInput("m", "Number of samples:", 2, min = 1, max = 100)
    ),
    mainPanel(
      plotOutput("hist")
    )
  )
)

server <- function(input, output, session) {
  output$hist <- renderPlot({
   means <- replicate(1e4, mean(runif(input$m)))
    hist(means, breaks = 20)
  }, res = 96)
}

shinyApp(ui, server)

```


## Multi-page layout
```{r}
ui <- fluidPage(
  tabsetPanel(
    tabPanel("Chennai Super Kings", 
      fileInput("file", "Data", buttonLabel = "Upload..."),
      textInput("delim", "Delimiter (leave blank to guess)", ""),
      numericInput("skip", "Rows to skip", 0, min = 0),
      numericInput("rows", "Rows to preview", 10, min = 1)
    ), # CSK tabPanel ends here
    tabPanel("Delhi",
             textInput("delim", "Delimiter (leave blank to guess)", "")
             ), # Delhi tabPanel ends here
    tabPanel("Mumbai Indians",
             numericInput("rows", "Rows to preview", 10, min = 1)
             ),
    tabPanel("Kings Punjab",
             numericInput("rows", "Rows to preview", 10, min = 1)
             ),
    tabPanel("RC Bangalore",
             numericInput("rows", "Rows to preview", 10, min = 1)
             ),
    tabPanel("Rajasthan Royals",
             numericInput("rows", "Rows to preview", 10, min = 1)
             ),
    tabPanel("Kolkata Knight Riders",
             numericInput("rows", "Rows to preview", 10, min = 1)
             ),
    tabPanel("Hyderabad Risers",
             numericInput("rows", "Rows to preview", 10, min = 1)
             )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```

## Multi-panels through tabs

```{r}
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      textOutput("panel")
    ),
    mainPanel(
      tabsetPanel(
        id = "tabset",
        tabPanel("panel 1", "one"),
        tabPanel("panel 2", "two"),
        tabPanel("panel 3", "three")
      )
    )
  )
)

server <- function(input, output, session) {
  output$panel <- renderText({
    paste("Current panel: ", input$tabset)
  })
}

shinyApp(ui, server)
```

## Navigation bars: Vertical
```{r}
ui <- fluidPage(
  navlistPanel(
    id = "tabset",
    "Heading 1",
    tabPanel("panel 1", "Panel one contents"),
    "Heading 2",
    tabPanel("panel 2", "Panel two contents"),
    tabPanel("panel 3", "Panel three contents")
  )
)

server <- function(input, output, session) {
  
}
shinyApp(ui, server)
```


## Navigation bars: Horizontal
```{r}
ui <- navbarPage(
  "Page title",   
  tabPanel("panel 1", "one"),
  tabPanel("panel 2", "two"),
  tabPanel("panel 3", "three"),
  navbarMenu("subpanels", 
    tabPanel("panel 4a", "four-a"),
    tabPanel("panel 4b", "four-b"),
    tabPanel("panel 4c", "four-c")
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


```{r}
ui <- navbarPage(
  "IPL Teams",   
  tabPanel("Chennai Kings", "one"),
  tabPanel("Delhi", "two"),
  tabPanel("Mumbai Indians", "three"),
  tabPanel("Kolkata Riders", "four"),
  tabPanel("RC Bangalore", "five"),
  tabPanel("Hyderabad Risers", "six"),
  tabPanel("Royals Rajasthan", "seven"),
  tabPanel("Kings Punjab", "eight"),
  navbarMenu("Year", 
    tabPanel("2020", "Twenty Twenty"),
    tabPanel("2019", "Twenty Nineteen"),
    tabPanel("2018", "Twenty Eighteen")
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```
## Themes
* Customise background and dashboard using support packages in R-Shiny ecosystem
* Refer to extensions of Shiny here: https://github.com/nanxstats/awesome-shiny-extensions

## Themes with bslib::bs_theme()
```{r}
library(shiny)

ui <- fluidPage(
  theme = bslib::bs_theme(bootswatch = "darkly"),
  sidebarLayout(
    sidebarPanel(
      textInput("txt", "Text input:", "text here"),
      sliderInput("slider", "Slider input:", 1, 100, 30)
    ),
    mainPanel(
      h1(paste0("Theme: darkly")),
      h2("Header 2"),
      p("Some text")
    )
  )
)


server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


## Blend theme with your panel
```{r}
library(ggplot2)

ui <- navbarPage(
  theme = bslib::bs_theme(bootswatch = "darkly"),
  "Page title",   
  tabPanel("panel 1", sidebarLayout(
    sidebarPanel(
      textInput("txt", "Text input:", "text here"),
      sliderInput("slider", "Slider input:", 1, 100, 30)
    ),
    mainPanel(
      h1(paste0("Theme: darkly")),
      h2("Header 2"),
      p("Some text")
    )
  )
           
           ),#tabpanel
  tabPanel("Dataset", 
           selectInput("dataset", 
                label = "Dataset BITS", 
                choices = ls("package:datasets")),
    verbatimTextOutput("summary"),
    tableOutput("table")
           ), #tablpanel
  tabPanel("Plot", 
            plotOutput("plot")
           
           ), #tabpanel
  navbarMenu("subpanels", 
    tabPanel("panel 4a", "four-a"),
    tabPanel("panel 4b", "four-b"),
    tabPanel("panel 4c", "four-c")
  )
)



server <- function(input, output, session) {
   # Create a reactive expression
    dataset <- reactive({
        get(input$dataset, "package:datasets")
    })
    
    output$summary <- renderPrint({
        # Use a reactive expression by calling it like a function
        summary(dataset())
    })
    
    output$table <- renderTable({
        dataset()
    })
    
     output$plot <- renderPlot({
    ggplot(mtcars, aes(wt, mpg)) +
      geom_point() +
      geom_smooth()
  }, res = 96)
     
}

shinyApp(ui, server)
```

## Interactive Graphics
* renderPlot()
* renderImage()
* Four mouse events
  * click, 
  * dblclick (double click), 
  * hover (when the mouse stays in the same place for a little while), and 
  * brush (a rectangular selection tool)
  
```{r}
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  verbatimTextOutput("info")
)

server <- function(input, output) {
  output$plot <- renderPlot({
    plot(mtcars$wt, mtcars$mpg)
  }, res = 96)

  output$info <- renderPrint({
    req(input$plot_click)
    x <- round(input$plot_click$x, 2)
    y <- round(input$plot_click$y, 2)
    cat("[", x, ", ", y, "]", sep = "")
  })
}

shinyApp(ui, server)
```

## Near points on a plot
```{r}
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  tableOutput("data")
)
server <- function(input, output, session) {
  output$plot <- renderPlot({
    plot(mtcars$wt, mtcars$mpg)
  }, res = 96)
  
  output$data <- renderTable({
    nearPoints(mtcars, input$plot_click, xvar = "wt", yvar = "mpg")
  })
}

shinyApp(ui, server)
```


## Near points using ggplot and browser
```{r}
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  tableOutput("data")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    ggplot(mtcars, aes(wt, mpg)) + geom_point()
  }, res = 96)
  
  output$data <- renderTable({
    req(input$plot_click)
     #browser()
    nearPoints(mtcars, input$plot_click)
  })
}

shinyApp(ui, server)
```

## Brushes or selection on Plots
```{r}
ui <- fluidPage(
  plotOutput("plot", brush = "plot_brush"),
  tableOutput("data")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    ggplot(mtcars, aes(wt, mpg)) + geom_point()
  }, res = 96)
  
  output$data <- renderTable({
    brushedPoints(mtcars, input$plot_brush)
  })
}

shinyApp(ui, server)
```


```{r}
set.seed(1014)
df <- data.frame(x = rnorm(100), y = rnorm(100))

ui <- fluidPage(
  plotOutput("plot", click = "plot_click", )
)

server <- function(input, output, session) {
  dist <- reactiveVal(rep(1, nrow(df)))
  observeEvent(input$plot_click,
    dist(nearPoints(df, input$plot_click, allRows = TRUE, addDist = TRUE)$dist_)  
  )
  
  output$plot <- renderPlot({
    df$dist <- dist()
    ggplot(df, aes(x, y, size = dist)) + 
      geom_point() + 
      scale_size_area(limits = c(0, 1000), max_size = 10, guide = NULL)
  }, res = 96)
}

shinyApp(ui, server)
```

## More interactions on graphs
```{r}
ui <- fluidPage(
  plotOutput("plot", brush = "plot_brush", dblclick = "plot_reset")
)
server <- function(input, output, session) {
  selected <- reactiveVal(rep(FALSE, nrow(mtcars)))

  observeEvent(input$plot_brush, {
    brushed <- brushedPoints(mtcars, input$plot_brush, allRows = TRUE)$selected_
    selected(brushed | selected())
  })
  observeEvent(input$plot_reset, {
    selected(rep(FALSE, nrow(mtcars)))
  })

  output$plot <- renderPlot({
    mtcars$sel <- selected()
    ggplot(mtcars, aes(wt, mpg)) + 
      geom_point(aes(colour = sel)) +
      scale_colour_discrete(limits = c("TRUE", "FALSE"))
  }, res = 96)
}

shinyApp(ui, server)
```

## Interactive UI for Graphs
```{r}
ui <- fluidPage(
  sliderInput("height", "height", min = 100, max = 500, value = 250),
  sliderInput("width", "width", min = 100, max = 500, value = 250),
  plotOutput("plot", width = 250, height = 250)
)
server <- function(input, output, session) {
  output$plot <- renderPlot(
    width = function() input$width,
    height = function() input$height,
    res = 96,
    {
      plot(rnorm(20), rnorm(20))
    }
  )
}

shinyApp(ui, server)
```

## Interactivity with Images
* Below illustration is working here: https://hadley.shinyapps.io/ms-puppies/
```{r}
puppies <- tibble::tribble(
  ~breed, ~ id, ~author, 
  "corgi", "eoqnr8ikwFE","alvannee",
  "labrador", "KCdYn0xu2fU", "shaneguymon",
  "spaniel", "TzjMd7i5WQI", "_redo_"
)

ui <- fluidPage(
  selectInput("id", "Pick a breed", choices = setNames(puppies$id, puppies$breed)),
  htmlOutput("source"),
  imageOutput("photo")
)
server <- function(input, output, session) {
  # output$photo <- renderImage({
  #   list(
  #     src = file.path("puppy-photos", paste0(input$id, ".jpg")),
  #     contentType = "image/jpeg",
  #     width = 500,
  #     height = 650
  #   )
  # }, deleteFile = FALSE)
  # 
  # output$source <- renderUI({
  #   info <- puppies[puppies$id == input$id, , drop = FALSE]
  #   HTML(glue::glue("<p>
  #     <a href='https://unsplash.com/photos/{info$id}'>original</a> by
  #     <a href='https://unsplash.com/@{info$author}'>{info$author}</a>
  #   </p>"))
  # })
}

shinyApp(ui, server)
```

## Download file
```{r}
ui <- fluidPage(
  selectInput("dataset", "Pick a dataset", ls("package:datasets")),
  tableOutput("preview"),
  downloadButton("download", "Download .tsv")
)

server <- function(input, output, session) {
  data <- reactive({
    out <- get(input$dataset, "package:datasets")
    if (!is.data.frame(out)) {
      validate(paste0("'", input$dataset, "' is not a data frame"))
    }
    out
  })
  
  output$preview <- renderTable({
    head(data())
  })
    
  output$download <- downloadHandler(
    filename = function() {
      paste0(input$dataset, ".tsv")
    },
    content = function(file) {
      vroom::vroom_write(data(), file)
    }
  )
}

shinyApp(ui, server)
```


## More interactive upload/download operations
* read file, clean it, and download in tsv format
```{r}
library(tidyverse)
library(janitor)

ui_upload <- sidebarLayout(
  sidebarPanel(
    fileInput("file", "Data", buttonLabel = "Upload..."),
    textInput("delim", "Delimiter (leave blank to guess)", ""),
    numericInput("skip", "Rows to skip", 0, min = 0),
    numericInput("rows", "Rows to preview", 10, min = 1)
  ),
  mainPanel(
    h3("Raw data"),
    tableOutput("preview1")
  )
)

ui_clean <- sidebarLayout(
  sidebarPanel(
    checkboxInput("snake", "Rename columns to snake case?"),
    checkboxInput("constant", "Remove constant columns?"),
    checkboxInput("empty", "Remove empty cols?")
  ),
  mainPanel(
    h3("Cleaner data"),
    tableOutput("preview2")
  )
)

ui_download <- fluidRow(
  column(width = 12, downloadButton("download", class = "btn-block"))
)

ui <- fluidPage(
  ui_upload,
  ui_clean,
  ui_download
)

server <- function(input, output, session) {
  # Upload ---------------------------------------------------------
  raw <- reactive({
    req(input$file)
    delim <- if (input$delim == "") NULL else input$delim
    vroom::vroom(input$file$datapath, delim = delim, skip = input$skip)
  })
  output$preview1 <- renderTable(head(raw(), input$rows))
  
  # Clean ----------------------------------------------------------
  tidied <- reactive({
    out <- raw()
    if (input$snake) {
      names(out) <- janitor::make_clean_names(names(out))
    }
    if (input$empty) {
      out <- janitor::remove_empty(out, "cols")
    }
    if (input$constant) {
      out <- janitor::remove_constant(out)
    }
    
    out
  })
  output$preview2 <- renderTable(head(tidied(), input$rows))
  
  # Download -------------------------------------------------------
  output$download <- downloadHandler(
    filename = function() {
      paste0(tools::file_path_sans_ext(input$file$name), ".tsv")
    },
    content = function(file) {
      vroom::vroom_write(tidied(), file)
    }
  )
}

shinyApp(ui, server)
```

