---
title: "Operations on Image-I"
author: "Revendranath T"
date: "28/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Image Processing and Operations in R
* magick is a R-package that binds to ImageMagick: https://imagemagick.org/index.php
* enables loading, editing, manipulating, and processing images
* supports image formats:  png, jpeg, tiff and more
* enables  manipulations such as rotate, scale, crop, trim, blur, flip, annotate and many more

```{r}
#install magick package from CRAN
#install.packages('magick') 
#install magick from Github ropensci library - note: requires RTools
#devtools::install_github('ropensci/magick')
```
## Load library and check which functions are supported by ImageMagick
```{r}
library(magick)
str(magick::magick_config())
```

```{r}
getwd()
```

## Read a image or images from the folders/internet using image_read()
```{r}
bits_open <- image_read('C:/Users/user/bits_open.jpg')
cat_saint <- image_read('C:/Users/user/cat_saint.jpg')
mac_cat <- image_read('C:/Users/user/mac_cat.png')
```

```{r}
bits_open %>% plot()
cat_saint %>% plot()
mac_cat %>% plot()
```

```{r}
#read a jpeg image 
bits <- image_read("https://www.bits-pilani.ac.in/uploads/Pilani/pilanimanagementadmin/Gallery/Untitled-3.jpg")

bits %>% plot()

#image of Maryam Mirzakhani
maryam <- image_read('https://images.theconversation.com/files/56344/original/ht2dfgcq-1407895853.jpg')

plot(maryam)
#or
maryam %>% plot()

bits_hyd <- image_read('https://universe.bits-pilani.ac.in/Uploads/Hyderabad/adminforbitshyderabad/Gallery/2019-3-30--10-36-51-730_54799191_2409526282393364_201582859813650432_n.jpg')

bits_hyd %>% plot()
```

## Basics of images
* image_info() provides basic details
```{r}
print(image_info(bits))
print(image_info(maryam))
print(image_info(bits_open))
print(image_info(cat_saint))
print(image_info(mac_cat))
print(image_info(bits_hyd))
```
## What does image_info provide?
* format 
* width
* height
* colorspace
* matte: Mattes are used in photography and special effects filmmaking to combine two or more image elements into a single, final image
* filesize:  0 because ImageMagick is lazy (in the good sense) and does not render until it has to.
* density

## Display image in R or RStudio 
```{r}
print(bits)
print(maryam)
#or
plot(bits)
plot(maryam)

typeof(maryam)

# Alternatively
# X11 only 
image_display(bits) #does not work as R is not run on x11

# System dependent
image_browse(maryam) # pops-up

```
## Convert image format
```{r}
image_info(bits)

bits %>% plot()
```
```{r}
bits_png <- image_convert(bits, "png")
image_info(bits_png)
```
```{r}
bits_png %>% plot()

image_info(bits_png)

image_info(bits_open)
```



## Write image to a folder 
```{r}
#getwd()
#C:\Users\user\Documents\R\R-Tutorial
# image_write(bits, path = 'C:/Users/user/Documents/R/R-Tutorial')
# 
# image_write(bits)

bits %>% 
  image_rotate(180) %>% 
image_write('bits_180.png','png')


```

## Image Enhancements
* Hue :Shift color of all pixels clockwise or counterclockwise about the color wheel.
* Saturation: Increase or decrease color saturation. 
  * Fully-saturated colors are very bright, while low saturation are grayish.
* Brightness/Lightness: 
  * Increase or decrease lightness. 
  * Zero lightness results in all black pixels, while full lightness is white.
* Use image_modulate()

```{r}
cat_saint %>% plot()
```


```{r}
# Brightness, Saturation, Hue
mod_img <- image_modulate(bits, 
                          brightness = 120, 
                          saturation = 20, 
                          hue = 20)

plot(mod_img)

image_modulate(cat_saint, 
               brightness = 220, 
               saturation = 70, 
               hue = 160) %>% 
  plot()
```
## Aspect ratio: 
* The aspect ratio of an image is the ratio of its width to its height. 
* It is commonly expressed as two numbers separated by a colon, as in 16:9. 
* For an x:y aspect ratio, the image is x units wide and y units high. 

## Geometry
* Values passed to image_scale() are called geometry
* Geometry provides a convenient means to specify a geometry argument. 
* Geometry specification use width, height, xoffset, and yoffset to specify size/dimension and position for an object and is represented in the form of:
* “<width>x<height>{+-}<xoffset>{+-}<yoffset>” ==> "300x400+50-75"
* x to represent the set the size. 
* Not mandatory to give both width and height 
  * instead just one of those is sufficient.
* 100 implies width = 100px, 
* image_scale(image, “100”) will resize the image proportionally to width: 100px
* And without the first number, x300 implies height = 300px, 
* so image_scale(image, “x300”) will resize the image proportionally to height: 300px
* xoffset and yoffset are used with either + or - to offset the image with respect to the left or top edges.
  * +xoffset : The left edge of the object is to be placed xoffset pixels in from the left edge of the image.
  * -xoffset : The left edge of the object is to be placed outside the image, xoffset pixels out from the left edge of the image.
  * +yoffset : The top edge of the object is to be yoffset pixels below the top edge of the image.
  * -yoffset : The top edge of the object is to be yoffset pixels above the top edge of the image.
  * like size, both xoffset and yoffset must be provided

## Image Transformation
* image editing application: cropping and scaling. 
* magick has two functions
  * image_crop() for cropping 
  *  image_scale() for scaling
  
```{r}
image_info(cat_saint)
```

```{r}
cat_saint %>% plot()
```
```{r}
# image_crop()
# without off-set
#with off-set
image_crop(cat_saint, "250x250+0+0") %>% plot()
```

```{r}
#with off-set
image_crop(cat_saint, "250x250+50+0") %>% plot()
```
```{r}
image_append(c(cat_saint,
              cat_saint %>% 
               image_crop("250x250+0+0"),
              cat_saint %>% 
               image_crop("250x250+50+0"),
              cat_saint %>% 
               image_crop("250x250+120+0")
              )
             ) %>% plot()
```

## image_scale()
* image_scale() acts similarly to image_crop() but scales down/up the entire image to the given size instead of cropping a part of it.
* image_scale and image_sample resize using simple ratio and pixel sampling algorithm.
```{r}
image_info(cat_saint)
```

```{r}

image_scale(cat_saint, "150x150") %>% plot()

image_scale(cat_saint, "150x150") %>% image_info()
```

```{r}
cat_saint %>% 
  image_scale("150x150") %>% 
  image_info()

image_append(c(cat_saint,
              cat_saint %>% 
               image_scale("150x150"),
               cat_saint %>% 
               image_scale("300"),
               cat_saint %>% 
              image_scale("x400"),
              cat_saint %>% 
              image_scale("x600")
              )
             ) %>% plot()

bits %>% 
  image_scale("400") %>% #width 400  
  plot()
  
bits %>% 
  image_scale("x400") %>% #heigth 400
  plot()

image_crop(bits, "x350") %>% plot()

image_crop(bits, "+150+50") %>% plot()

#what's the difference?


```
## More examples
```{r}
bits %>% 
  image_crop("100x150+50")

#rotate image
# image_rotate rotates and increases size of canvas to fit rotated image.
bits %>% 
  image_rotate(90) %>% 
  image_rotate(-90) %>% 
  plot()
```

```{r}
# image_flip and image_flop invert image vertically and horizontally
# image flip
cat_saint %>% plot()
  
cat_saint %>% 
  image_flip() %>% 
  plot()
```

```{r}
#image flop
image_append( c(cat_saint,
                cat_saint %>% 
                  image_flip(),
                cat_saint %>% 
                  image_flop() )
              ) %>% 
    plot()
```

```{r}

image_append( c(bits,
                bits %>% 
                  image_flop() )
              ) %>% 
    plot()

#fill image with colors
bits %>% 
image_fill("orange", 
           point = "+100+300", 
           fuzz = 30
           ) %>% 
  plot()

```

```{r}
bits %>% plot()
```

```{r}
bits %>% image_flop( ) %>% plot()
```


# Filters and effects on images
```{r}
#use image_blur to introduce blur to images
bits %>% 
  image_blur(10, 5) %>% 
  plot()

cat_saint %>% 
  image_blur(10, 5) %>% 
  plot()
```

```{r}
#add some noise to images
bits %>% 
  image_noise() %>% 
  plot()

cat_saint %>% 
  image_noise() %>% 
  plot()
```

```{r}
#would you like to do a pencil drawing?
#charcoal filter (pencil drawing)
bits %>% 
  image_charcoal() %>%
  plot()

cat_saint %>% 
  image_charcoal() %>%
  plot()

# How about an oil paint?
bits %>% 
  image_oilpaint() %>%
  plot()
# didn't come so well, does it? Try with another image

maryam %>% 
  image_oilpaint() %>%
  plot()

#negate filter
image_append( c( bits,
                bits %>% 
                 image_charcoal(),
               bits %>% 
                 image_negate()
               ) 
             ) %>%
  plot()

bits %>% plot()

bits %>% image_negate() %>% plot()

bits %>% image_negate() %>% image_negate() %>% plot()

bits %>% image_charcoal() %>% image_negate() %>% plot()
```
```{r}
bits %>%  plot()
```


# Other basic transformations on images
```{r}

# image_trim removes edges that are the background color from the image.
image_trim(bits)
# 

image_append( c(bits,
                bits %>% 
  image_border(color="red") %>% 
  image_trim(fuzz = 10 ),
  
   bits %>% 
   image_border(color="red") %>% 
  image_trim(fuzz = 40 )
                
                )) %>% 
  plot()

bits %>% 
  image_border(color="red") %>% 
  image_trim(fuzz = 10 )
  plot() 

# image_crop cuts out a subregion of original image
#image_info(bits) %>%
  
  bits %>% 
  image_crop( "600x300+50") %>% 
  plot()
  
  image_info(cat_saint)
  
  cat_saint %>% 
  image_crop( "600x300") %>% 
  plot()

# image_chop removes vertical or horizontal subregion of image.

image_append( c(bits, bits %>% 
  image_chop("100x20"))) %>% 
  plot()
```

```{r}


# image_scale() and image_sample() resize using simple ratio and pixel sampling algorithm

image_info(bits)

image_append(c(bits, bits %>% 
  image_sample("300x300"),
  bits %>% image_scale("300x300")
  )) %>% 
  plot()

# The most powerful resize function is image_resize which allows for setting a custom resize filter.
# Output of image_scale is similar to image_resize(img,filter = "point").
# For resize operations it holds that if no geometry is specified, all frames are rescaled to match the top frame.
# image_resize resizes using custom filterType

image_append(c(bits,
bits %>% 
  image_resize("300x300+50", filter = NULL))) %>% 
  plot()

# image_deskew auto rotate to correct skewed images
image_append(c(bits,bits %>%  
  image_deskew( threshold = 90))) %>% 
  plot()

image_append(c(cat_saint,cat_saint %>%  
  image_deskew( threshold = 90))) %>% 
  plot()

image_append(c(mac_cat,mac_cat %>%  
  image_deskew( threshold = 90))) %>% 
  plot()

image_append(c(bits, bits %>% 
  image_page(pagesize = NULL, density = NULL),
  bits %>% 
  image_page(pagesize = 1, density = 72)
  )) %>% 
  plot()


image_append(c(mac_cat, mac_cat %>% 
  image_page(pagesize = NULL, density = NULL),
  mac_cat %>% 
  image_page(pagesize = 1, density = 72)
  )) %>% 
  plot()

```
```{r}
bits %>%
    image_crop('x420') %>%
    image_negate() %>%
    plot()

cat_saint %>%
    image_crop('x420') %>%
    image_negate() %>%
    plot()
```

##Gray scaling image
```{r}
# crop out width:100px and height:150px starting offset +50px from the left
bits %>% 
  image_quantize(colorspace = 'gray') %>% 
  plot()

#resize proportionally to width: 200px
bits %>% 
  image_scale(200)

# flood fill with blue starting at the point at x:100, y:200
bits %>% 
  image_fill("blue", "100x100") %>% 
  plot()

image_info(cat_saint)

 
  image_append( c(cat_saint, 
                  cat_saint %>% image_fill("blue", "10x10"))) %>% 
    plot()

#adds a border of 20px left+right and 10px top+bottom
image_append(c(bits, bits %>% 
  image_border( "red", "20x10"))) %>% 
  plot()
```
## Kernal convolution
* Kernal: in image processing
  * A kernel, convolution matrix, or mask is a small matrix. 
  * It is used for blurring, sharpening, embossing, edge detection, and more. 
  * Accomplished by doing a convolution between a kernel and an image.

```{r}
kernal <- matrix(0, ncol = 3, nrow = 3)
kernal
kernal[1, 2] <- 0.25
kernal
kernal[2, c(1, 3)] <- 0.35
kernal
kernal[3, 2] <- 0.45
kernal
```
```{r}
kernal1 <- matrix(0, ncol = 3, nrow = 3)

kernal1[1,c(1,3)] <- 0.25 
kernal1[3,c(1,3)] <- 1.25 
kernal1
```

```{r}
bits %>% 
  plot()
```


## Kernal in practice
```{r}
kernal

bits %>% 
  image_convolve(kernal) %>%
  plot()

image_append(c(bits, 
               bits %>% 
                 image_convolve(kernal),
               bits %>% 
                 image_convolve(kernal1)
               )
             ) %>% plot()
```

## Negate and check the kernal effect
```{r}
image_append(c(bits %>% image_negate(), 
               bits %>% 
                 image_convolve(kernal) %>% image_negate(),
               bits %>% 
                 image_convolve(kernal1) %>% image_negate()
               )
             ) %>% plot()

image_append(c(bits, 
               bits %>% 
                 image_convolve(kernal),
               bits %>% 
                 image_convolve(kernal1)
               )
             ) %>% plot()
```
# Standard Kernals
* Blurring Kernels
  * Unity or Identity Kernel
  * Shaped Mean or Average Kernel
  * Gaussian Blurring Kernel
  * Linear 1D Gaussian (Blur)
  * Half Gaussian (Comet)
* Edge Detection Kernels
  * Laplacian Of Gaussians (LoG)
  * Difference Of Gaussians (DoG)
  * Discrete Laplacian Kernels
* Directional Kernels
  * Sobel,  
  * Roberts,  
  * Prewitt,  
  * Compass,  
  * Kirsch,  
  * FreiChen

```{r}
bits %>% 
  image_convolve('FreiChen') %>%  # Blur Comet LoG DoG Laplacian Roberts Prewitt Compass Kirsch FreiChen
  image_negate() %>% 
  plot()


image_append(c(  bits %>% 
                 image_convolve('FreiChen'),
               bits %>% 
                 image_convolve('Roberts')
               )
             ) %>% plot()

```

## Text Annotation
* Add text with images
```{r}
bits %>% image_annotate("Welcome to BITS Pilani!", 
                        size = 70, 
                        boxcolor = "pink",
                        gravity = "southwest", 
                        color = "red") %>% 
  plot()

bits %>% 
  image_annotate( "CONFIDENTIAL", 
                         size = 30, 
                         color = "red", 
                         #boxcolor = "pink",
                         degrees = -45, 
                         location = "+450+200") %>% 
  plot()

#Supported fonts: "sans", "mono", "serif", "Times", "Helvetica", "Trebuchet", "Georgia", "Palatino"or "Comic Sans"
bits %>%
  image_annotate("Welcome to BITS Pilani!", 
                 font = 'Times', 
                 size = 30) %>% 
  plot()

```

```{r}
bits %>%
  image_rotate(270) %>%
  image_background("blue", flatten = TRUE) %>%
  image_border("red", "10x10") %>% 
  image_annotate("This is BITS Pilani", color = "red", size = 70,
                 location = "+300+200", degrees = 45) %>% 
  plot()
```
## Image vectors
* Functions in magick have been vectorized to support working with layers, compositions or animation.
* The standard base methods [ [[, c() and length() are used to manipulate vectors of images which can then be treated as layers or frames.
```{r}
veri <- image_read('C:/Users/user/images/veretasium.gif')

veri

veri %>% plot()

image_info(veri)

#what did you observe in the above output?

length(veri)

veri %>% 
 image_scale("200x") %>%
  image_quantize(128)
```

## Flip the gif
```{r}
rev(veri) %>% 
 # image_flip() %>% 
  image_annotate("Reverse", size = 20, color = "white")

## how to change it
veri %>% 
  image_rotate(180) %>% 
  rev() %>% 
  image_flip() %>% 
  image_annotate("Reverse", size = 20, color = "white")

```

## Layers of Images
```{r}
img_comb <- c(image_scale(bits, "300x300"), 
                          image_scale(cat_saint, "200x200"), 
                          image_scale(mac_cat, "100x100+50+50"))

c(bits, cat_saint, mac_cat) %>% image_info()

class(img_comb)

image_info(img_comb)

#img_comb  <-  image_scale(img_comb, "100x100") 

image_mosaic(img_comb) %>% plot()

image_flatten(img_comb) %>% plot()


```

### Mosaic and Flattening has some composite operators
* More details here: https://www.imagemagick.org/Magick++/Enumerations.html#CompositeOperator

* Some examples below
```{r}
image_flatten(img_comb, 'Add')
#or
img_comb %>% 
image_flatten( 'Add') %>% 
  plot()

img_comb %>% 
image_flatten('Modulate') %>% 
  plot()

img_comb %>% 
image_flatten('Minus') %>% 
  plot()


```
## Image Composing
```{r}
 cat_saint %>% 
  image_scale("x400") %>% 
  image_rotate(300) %>% 
  image_background("red") %>% 
  image_composite(cat_saint,
                  offset = "+180+100")
  

```


```{r}
## Another illustration
bigdata <- image_read('https://jeroen.github.io/images/bigdata.jpg')
frink <- image_read("https://jeroen.github.io/images/frink.png")
logo <- image_read("https://jeroen.github.io/images/Rlogo.png")
img <- c(bigdata, logo, frink)
image_info(img)
image_mosaic(img) 
image_flatten(img)
image_flatten(img, 'Add')
image_flatten(img, 'Modulate')
image_flatten(img, 'Minus')

#Combing
image_append(image_scale(img, "x200"))
image_append(image_scale(img, "100"), stack = TRUE)

bigdatafrink <- image_scale(image_rotate(image_background(frink, "none"), 300), "x200")
image_composite(image_scale(bigdata, "x400"), bigdatafrink, offset = "+180+100")

#Reference: https://cran.r-project.org/web/packages/magick/vignettes/intro.html
```

### Animation
```{r}

img_comb %>% 
  image_scale("200x200") %>% 
  image_animate(fps = 1, dispose = "previous")

#BITS Logos

bits1 <- image_read('C:/Users/user/images/bits_logos/BITS_Pilani-Logo.png')
bits2 <- image_read('C:/Users/user/images/bits_logos/bits2.jpg')
bits3 <- image_read('C:/Users/user/images/bits_logos/bits3.png')


c(bits1, bits2, bits3) %>% 
  image_scale("300x300") %>% 
  image_animate(fps = 2, dispose = "previous")
  
```

## Text Extraction from PDFs
* While reading a PDF document, each page is converted to an element of the vector.
* Specify the density while reading PDF, as PDF is rendered while reading.
```{r}
# Objective is to read this pdf document: https://www.bits-pilani.ac.in/downloads/recruitment.pdf

recruitment <-  image_read_pdf('https://www.bits-pilani.ac.in/downloads/recruitment.pdf', density = 72)

class(recruitment)

image_info(recruitment)

#display the first page
recruitment[1]

recruitment[15]
```

## Text Extraction from Images
* Otherwise called OCR text extraction
* Extract text from images using OCR. 
* Use the tesseract package:
```{r}
#install.packages("tesseract")

library(tesseract)

image_text <- image_read('C:/Users/user/images/text_data/image_text.png')

class(image_text)

#check the image
print(image_text)

image_text %>% plot()

#Extract test
cat(image_ocr(image_text))
#or
image_text %>% 
  image_ocr() %>% 
  cat

recruitment[2] %>% 
  image_ocr() 
```

## Combine image and pdf to extract text from pdf
```{r}
recruitment[1] %>% 
image_ocr() %>% cat

recruitment[3] %>% 
  image_ocr() %>% 
  cat
```


## References
* https://towardsdatascience.com/advanced-image-processing-in-r-210618ab128a
* https://heartbeat.fritz.ai/image-manipulation-for-machine-learning-in-r-ff2b92069fef
* https://cran.r-project.org/web/packages/imager/vignettes/gettingstarted.html
* https://cran.r-project.org/web/packages/magick/vignettes/intro.html#Build_from_source
