---
title: "joins in tidyverse"
author: "Revendranath T"
date: "27/9/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

## joins
* mutating
* filtering
* nesting

## Mutating joins: combine variables from the two data.frames:
```{r}
print(band_members)
print(band_instruments)
```


### inner_join()
* return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned.

```{r}
band_members %>% inner_join(band_instruments, by="name")
```

### left_join() 
* return all rows from x, and all columns from x and y. Rows in x with no match in y will have NA values in the new columns. If there are multiple matches between x and y, all combinations of the matches are returned.
```{r}
band_members %>% left_join(band_instruments, by ="name")
#or
 band_instruments %>% right_join(band_members, by ="name")
#india %>% left_join(usa, by ="product_id")
```

### right_join()
* return all rows from y, and all columns from x and y. Rows in y with no match in x will have NA values in the new columns. If there are multiple matches between x and y, all combinations of the matches are returned.

```{r}
band_members %>% right_join(band_instruments, by ="name")
#or
band_instruments %>% left_join(band_members, by ="name")
```

### full_join()
* return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing.
```{r}
band_members %>% full_join(band_instruments, by ="name")
```

```{r}
band_instruments %>% full_join(band_members, by ="name")
```

## Filtering joins keep cases from the left-hand data.frame:

### semi_join()
* return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x.
```{r}
band_members %>% inner_join(band_instruments, by = "name") %>% 
  select(name, band)

band_members %>% semi_join(band_instruments, by = "name")

band_instruments %>% semi_join(band_members, by = "name")
```

### anti_join()
* return all rows from x where there are not matching values in y, keeping just columns from x.
```{r}
print(band_members)
print(band_instruments)
```



```{r}
#print(band_members %>% inner_join(band_instruments, by = "name"))
print(band_members %>% anti_join(band_instruments, by = "name"))
```

## Nesting joins create a list column of data.frames:
* "Nesting" joins keep cases from the LHS and nests the RHS

### nest_join()
* return all rows and all columns from x. Adds a list column of tibbles. 
* Each tibble contains all the rows from y that match that row of x. When there is no match, the list column is a 0-row tibble with the same column names and types as y.
* nest_join() is the most fundamental join since you can recreate the other joins from it. 
* An inner_join() is a nest_join() plus an tidyr::unnest(), and left_join() is a nest_join() plus an unnest(.drop = FALSE). 
* A semi_join() is a nest_join() plus a filter() where you check that every element of data has at least one row, and 
* an anti_join() is a nest_join() plus a filter() where you check every element has zero rows.
```{r}
# "Nesting" joins keep cases from the LHS and nests the RHS

band_members %>% nest_join(band_instruments, by = "name") %>% select(band_instruments) %>% glimpse()

print(band_instruments %>% nest_join(band_members, by = "name"))
```

```{r}
# To suppress the message, supply by
band_members %>% inner_join(band_instruments, by = "name")
# This is good practice in production code

```

```{r}
# Use a named `by` if the join variables have different names
band_members %>% full_join(band_instruments2, by = c("name" = "artist"))

# Note that only the key from the LHS is kept
```


## References
* https://dplyr.tidyverse.org/reference/join.html
* https://www.r-bloggers.com/2018/10/how-to-perform-merges-joins-on-two-or-more-data-frames-with-base-r-tidyverse-and-data-table/
