---
title: "Data Transformation (Tidyverse)"
author: "Revendranath T"
date: "9/20/2021"
output:
  slidy_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
#load libary

library(tidyverse)

knitr::opts_chunk$set(echo = FALSE)
```

# Goals and Introduction

By the end of this tutorial, you will know:

+ What "tidy data" is and why it's an awesome format
+ How to do some stuff with tidy data
+ How to get your data to be tidy
+ Some tips'n'tricks for dealing with "medium data" in R
+ functions & pipes (%>%) are at the heart of tidyverse
+ pipe (%>%) shortcut Ctrl+Shift+M
Uses the famous `iris` dataset that comes with `ggplot2` package


## iris dataset
* Details about the dataset: https://www.kaggle.com/uciml/iris
* Further details: https://archive.ics.uci.edu/ml/datasets/iris

```{r}
View(iris)
head(iris)
str(iris)
iris %>% glimpse()
summary(iris)
#or
iris %>% summary() %>% knitr::kable()

mean(iris$Petal.Length) #readable
#or
iris$Petal.Length %>% 
  mean() #readable?

#instead we can use following
mean_PL <- iris$Petal.Length %>% mean()

# mean for unique values
mean(unique(iris$Petal.Length))
#or
iris$Petal.Length %>% unique() %>% mean(na.rm=TRUE)

iris$Petal.Length %>% unique() %>% mean()

#round mean for unique values
round(mean(unique(iris$Petal.Length, na.rm=TRUE)), digits = 2)
#or

iris$Petal.Length %>% 
  unique() %>% 
  mean() %>% 
  round(digits = 2)

#debug error
iris%>% 
  select(Petal.Length) %>% 
  unique %>% 
  mean(na.rm = TRUE) %>% 
  round(digits = 2)

iris$Petal.Length #base R

iris %>% 
  select(Petal.Length) #Tidyverse

iris %>% 
  select(-c(Petal.Length, 
           Petal.Width)
         )
```
```{r}
iris %>% str()
```

## NYC Flights Dataset
* On-time data for a random sample of flights departing New York City airports in 2013.
```{r}
library(nycflights13)
library(tidyverse)

str(flights)

flights %>% head(5)

flights %>% tail()
```
### data types in the dataset flights
* int stands for integers.
* dbl stands for doubles, or real numbers.
* chr stands for character vectors, or strings.
* dttm stands for date-times (a date + a time).
* lgl stands for logical, vectors that contain only TRUE or FALSE.
* fctr stands for factors, which R uses to represent categorical variables with fixed possible values.
* date stands for dates.


### Basic transformation operations
* Pick observations by their values (filter()).
* Reorder the rows (arrange()).
* Pick variables by their names (select()).
* Create new variables with functions of existing variables (mutate()).
* Collapse many values down to a single summary (summarise())
* To group use: group_by()

### FILTER
```{r}
filter(flights, month == 1, day == 1) %>% 
  select(flight, carrier, dep_delay ) %>% 
  group_by(carrier) %>% 
  summarise( mean = mean(dep_delay))

flights %>%  
  filter( month == 1, day == 1) %>% 
  select(flight, carrier, dep_delay ) %>% 
  group_by(carrier) %>% 
  summarise( mean_name = mean(dep_delay), 
             median_name = median(dep_delay), 
             sd_name = sd(dep_delay))


# Assign using filter

jan1 <- filter(flights, month == 1, day == 1, year==2013)
#or
jan1 <- flights %>% filter(month == 1, 
                           day == 1, 
                           year==2013)
jan1

# Assign & display

(dec25 <- filter(flights, month == 12, day == 25))

(dec25 <- flights %>% filter(month == 12, 
                             day == 25)
  )

```

#### Comparison using FILTER
* To use filtering effectively, you have to know how to select the observations that you want using the comparison operators. 
* R provides the standard suite: >, >=, <, <=, != (not equal), and == (equal).

```{r}
# a = 1
# a <- 1
#Is this right?
filter(flights, month = 1) 
#or is this?
filter(flights, month == 1)

#The following code finds all flights that departed in November or December:
filter(flights, month == 11 | month == 12)

flights %>% 
  filter( month == 11 | 
            month == 12 |
            month == 1)

# Is this right to find all flights departed in November or December?
filter(flights, month == (11 | 12))
flights %>% 
  filter( month == (11 |  12))
# Dig what is wrong in the above code

x <- 12

x
x == (11||12)

x <- 7
(x == 11 | x ==12)
```
#### use %in% to simplify FILTER
```{r}
#Flight details in the month of november and december and jan
nov_dec_jan <- filter(flights, month %in% c(11, 12,1))
#or
nov_dec_jan <- flights %>% 
  filter(month %in% c(11, 12,1))

nov_dec_jan %>% filter( month == 12)
```


### Arrange values
* arrange() works similarly to filter() 
* except that instead of selecting rows, it changes their order. 
* It takes a data frame and a set of column names (or more complicated expressions) to order by. 
* If you provide more than one column name, each additional column will be used to break ties in the values of preceding columns
```{r}
arrange(flights, year, month, day)

flights %>% arrange(year, 
                    month, 
                    day)
# What is the order by default?

#arrange(flights, desc(year, month, day))
flights %>%  arrange(desc(month, 
                          day, 
                          year)
                     ) 

flights$year %>% unique

#Use desc() to re-order by a column in descending order:
arrange(flights, desc(dep_delay))

flights %>% arrange(desc(dep_delay)
                    )

flights %>%  
  filter( month == 1, day == 1) %>% 
  select(flight, carrier, dep_delay ) %>% 
  group_by(carrier) %>% 
  summarise( mean_name = mean(dep_delay), 
             median_name = median(dep_delay), 
             sd_name = sd(dep_delay)) %>% 
  arrange( desc(mean_name)
  )

#Missing values come at the end
df <- tibble(x = c(5, NA, 2, NA))
df
#df <- data.frame(x = c(5, 2, NA))

arrange(df, x)
df %>%  arrange(x)

arrange(df, desc(x))

df %>%  arrange(desc(x))

```

### Select columns with select()
```{r}
flights %>% glimpse()

flights[,c(1,7,10)]

# Select columns by name
select(flights, year, month, day)
#or
flights %>% 
  select(year, month, day)

flights %>% 
  select(flight, carrier, dep_delay)

# Select all columns between year and day (inclusive)
select(flights, year:dep_time)
#or
flights %>% select(year:dep_time)

# Select all columns except those from year to day (inclusive)
select(flights, -(year:dep_time))
#or
flights %>% select(-(year:dep_time))

```

### Helper functions in select
* starts_with("abc"): matches names that begin with “abc”.
* ends_with("xyz"): matches names that end with “xyz”.
* contains("ijk"): matches names that contain “ijk”.
* matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings. matches("f2018")
* num_range("x", 1:3): matches x1, x2 and x3
* To rename variables, do not use select(), use rename() which is a variant of select() that keeps all the variables that aren’t explicitly mentioned
* select() in conjunction with the everything() helper. This is useful if you have a handful of variables you’d like to move to the start of the data frame.

```{r}
flights %>% glimpse()
rename(flights, tail_num = tailnum) 
#or
flights %>% rename(tail_num = tailnum)  %>% glimpse()

#(flights <- flights %>% rename(tail_num = tail_num))

select(flights, time_hour, air_time, everything())
#or
flights %>% select( time_hour, 
                    air_time, 
                    everything())

flights %>% select(starts_with("dep_"), 
                   ends_with("t"),
                   contains("rr"))
```

## Add new variables with mutate()
* mutate() always adds new columns at the end of your dataset so we’ll start by creating a narrower dataset so we can see the new variables.
```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
# or

flights_sml <- flights %>%  
  select( 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

flights_sml %>% glimpse()

flights_sml %>% filter(dep_delay <0) 

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)

flights_sml %>% 
  filter(dep_delay > 0) %>% 
  mutate(gain = dep_delay - arr_delay,
  speed = (distance / air_time) * 60
)

flights_sml %>% glimpse()


flights_sml <- flights_sml %>% 
  filter(dep_delay > 0) %>% 
  mutate(gain = dep_delay - arr_delay,
  speed = (distance / air_time) * 60
)

flights_sml %>% glimpse()

#Once new variables are created they can be referred and used to create new variables
#The code below will throw error: Why?
flights_sml %>% mutate( hours = air_time/60,
                        gain_per_hour = gain/hours,
                        gain = dep_delay-arr_delay)

#the code below works to create new variables, why?
mutate(flights_sml,
  gain = dep_delay-arr_delay,
  hours = air_time/60,
  gain_per_hour = gain/hours
)
# or
flights_sml %>% mutate( gain = dep_delay-arr_delay,
                        hours = air_time/60,
                        gain_per_hour = gain/hours)

# If you only want to keep the new variables, use transmute():



transmute(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

#or
flights_sml %>% transmute(gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)

## throws an error below:
flights_sml %>% transmute( gain_per_hour = gain / hours,
  hours = air_time / 60,
  gain = dep_delay - arr_delay
)

flights_sml %>% transmute(hours = air_time / 60,
                          gain = dep_delay - arr_delay,
                          gain_per_hour = gain / hours
                          )
```

### Key considerations while creating variables
*  mutate() or transmute() must take a vector of values as input, return a vector with the same number of values as output.
* Examples: arithmetic operations, logs, offsets and leads, logical operations, cumulative and rolling aggregates,  ranking, etc.

```{r}
x <- flights$dep_time
class(x)
x
```


```{r}
#modular arithmetic
transmute(flights,
  dep_time,
  hour = dep_time %/% 100, #integer division
  minute = dep_time %% 100 # reminder 
)
#or
flights %>% 
  transmute( dep_time,
  hour = dep_time %/% 100, #integer division
  minute = dep_time %% 100 # reminder 
)

#examples of lag & lead
(x <- 1:10)
print(x)
print(lag(x))
print(lead(x))

# cumulative aggregates
print(x)
print(cumsum(x))
print(cummean(x))

#ranking 
y <- c(1, 2, 2, NA, 3, 4)
print(y)
print(min_rank(y))
print(desc(y))

print(min_rank(desc(y)))

print(sort(y))
print(min_rank(sort(y)))
print(sort(desc(y)))
print(min_rank(sort(desc(y))))

#If min_rank() doesn’t do what you need, look at the variants row_number(), dense_rank(), percent_rank(), cume_dist(), ntile().
```
```{r}
summary(y)
```

## Grouped summaries with summarise()
* summarise() provides summary & collapses a data frame to a single row

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
#or
flights %>% summarise(mean_delay = mean(dep_delay, 
                                   na.rm = TRUE),
                      median_delay = median(dep_delay, 
                                   na.rm = TRUE),
                      rows= n()
                      )
```
* When you need to summarise by groups then use group_by()
```{r}
by_day <- group_by(flights, year, month, day)
#or
by_day <- flights %>%  
  group_by(year, 
           month, 
           day)

summarise(by_day, delay = mean(dep_delay, na.rm = TRUE))

```

## Power of tidyverse in R through piping
* Combining multiple operations with the pipe

```{r}
by_dest <- group_by(flights, dest)
#or
by_dest <- flights %>% 
  group_by(dest)

delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delay

delays <- filter(delay, count > 20, dest != "HNL")

delay %>% filter(count > 10000,
                 dest != "HNL")

# Perform above operations using piping
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")

#Plot the same. We will do plotting in the 2nd part of this session
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
#or
delay %>% ggplot(mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```
### Other operations
* Missing values
* Counts: Whenever you do any aggregation, it’s always a good idea to include either a count (n()), or a count of non-missing values (sum(!is.na(x))). That way you can check that you’re not drawing conclusions based on very small amounts of data.
```{r}
x <- c(1, NA, 3)
print(x)
print(is.na(x))
print(!is.na(x))
#missing values
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

#Compare without missing values
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))

#What does this do?
not_cancelled <- flights %>%
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

#COUNTS
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay, na.rm = TRUE),
    n = n()
  ) 
delays %>% glimpse()
View(delays)
head(delays)

```

### Useful summary functions
```{r}
# Measures of location:  mean(x), but median(x)
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    avg_delay_all = mean(arr_delay),
    avg_delay_late = mean(arr_delay[arr_delay > 0]),# the average positive delay
    avg_delay_early = mean(arr_delay[arr_delay < 0]),# the average negative delay
  )

# Why is distance to some destinations more variable than to others?
#Measures of position: first(x), nth(x, 2), last(x)
not_cancelled %>% 
  group_by(dest) %>% 
  summarise(distance_sd = sd(distance)) %>% 
  arrange(desc(distance_sd))

# When do the first and last flights leave each day?
#Measures of position: first(x), nth(x, 2), last(x)
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first = min(dep_time),
    last = max(dep_time)
  )

# Measures of position: first(x), nth(x, 2), last(x)
not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(
    first_dep = first(dep_time), 
    last_dep = last(dep_time)
  )

```

### Group by multiple variables & ungrouping
```{r}
daily <- group_by(flights, year, month, day)
#or
daily <- flights %>% 
  group_by(year, month, day)

daily

(per_day   <- summarise(daily, flights = n()))

(per_month <- summarise(per_day, flights = sum(flights)))

(per_year  <- summarise(per_month, flights = sum(flights)))

#or
flights %>% 
  select(year, month, day) %>% 
  summarise(flights_counts =n()
            )
#ungrouping
#If you need to remove grouping, and return to operations on ungrouped data, use ungroup()
daily %>% 
  ungroup() %>%             # no longer grouped by date
  summarise(flights = n())  # all flights

```

### Grouped mutates (and filters)
```{r}
#Find the worst members of each group:
flights_sml %>% 
  group_by(year, month, day) %>%
  filter(rank(desc(arr_delay)) < 10)

#Find all groups bigger than a threshold
popular_dests <- flights %>% 
  group_by(dest) %>% 
  filter(n() > 365)
popular_dests

#Standardise to compute per group metrics:
popular_dests %>% 
  filter(arr_delay > 0) %>% 
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>% 
  select(year:day, dest, arr_delay, prop_delay)
```

