---
title: "Iterations - purrr package"
author: "Revendranath T"
date: "8/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# attach tidyverse meta-packages
```{r}
library(tidyverse)
```

```{r}
# defining a function which returns square
square <- function(x){
  return(x*x)
}

# Create a vector of number
vector1 <- c(2,4,5,6)

square(vector1)
```

```{r}
list1 <- list(c(2,4,5,6), c(7, 8, 9))
list1
```

```{r}
square(list1) # throws error
square(list1[[1]])
square(list1[[2]])
```

```{r}
df1 <- data.frame(c(2,4,5,6), 
                  c(7, 8, 9, 10),
                  c(2,4,5,6), 
                  c(7, 8, 9, 10),
                  c(2,4,5,6))

df1
```

```{r}
square(df1)
```
# Can you pass data frame to calculate mean?
```{r}
mean(df1)
```


## Calculate mean of five columns in a dataframe.
```{r}
## create sample dataframe

df <- tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10),
  e = rnorm(10)
)
df

square(df)
```

# Method-1 : manual calculation of means & medians of five columns separately
```{r}
mean(df$a)
median(df$a)
mean(df$b)
median(df$b)
mean(df$c)
median(df$c)
mean(df$d)
median(df$d)
mean(df$e)
mean(df$e)
```
```{r}
df[[3]]
```

## Method-2: write for loop

```{r}
output_mean <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output_mean[[i]] <- mean(df[[i]])      # 3. body
}
output_mean

output_median <- vector("double", ncol(df))  # 1. output
for (i in seq_along(df)) {            # 2. sequence
  output_median[[i]] <- median(df[[i]])      # 3. body
}
output_median

```


## The map functions in purrr package
* map() makes a list.
* map_lgl() makes a logical vector.
* map_int() makes an integer vector.
* map_dbl() makes a double vector.
* map_chr() makes a character vector.

** Each function takes a vector as input, applies a function to each piece, and then returns a new vector that’s the same length (and has the same names) as the input. 
** The type of the vector is determined by the suffix to the map function.

## We can use map functions to use same calculations as loops do
```{r}
map_dbl(df, mean)
#or 
df %>% map_dbl(mean)

df %>% map_dbl(median)
```

```{r}

df %>% select(a,b) %>% mean() # throws an error

df %>% select(a,b) %>% 
  summarise(mean_a = mean(a),
            mean_b = mean(b))

df %>% select(a,b) %>%  map_dbl(mean)
```

```{r}
map_dbl(df, median)
#or
df %>% map_dbl(median)

map_dbl(df, sd)
#or

df %>% map_dbl(sd) %>% typeof()



```
```{r}

z <- list(x = 1:3, y = 4:5)

map_int(z, length) %>% typeof()
```

## Shotcuts
* using split()
* Fit a linear model to each group in a dataset. 
* Example splits mtcars dataset into three pieces (one for each value of cylinder) and fits linear regression model to each piece:
```{r}
model <- lm(mpg ~ wt, data = mtcars)
model
summary(model)
```

```{r}
models <- mtcars %>% 
  split(.$cyl) %>% 
  map(function(df) lm(mpg ~ wt, data = df))

lm_function <- function(df)
{
  lm(mpg ~ wt, data = df)
}

models2 <- mtcars %>% 
  split(.$cyl) %>% 
  map(lm_function)
 
 
models

map(models, summary)

map(models2, summary)

models %>% 
  map(summary) %>% 
  map_dbl(~.$r.squared)
```
## Errors handling after iterations
* safely(): returns two elements
  * result is the original result. If there was an error, this will be NULL.
  * error is an error object. If the operation was successful, this will be NULL.
* safely() is designed to work with map
```{r}


y <- log(10)
y
y <- log("10")
y
```

```{r}
safe_log <- safely(log)
x <- safe_log(10)
x
x$result
x$error
```

```{r}
x
x$result

log("a")
```

```{r}
x <- safe_log("a")
x
```

## safely() working with map
```{r}
x <- list(1, 10, "a")
log("a")

y <- x %>% map(safely(log)) %>% str()

y[[3]]
```


## Mapping over multiple arguments
*use pmap()
```{r}
mu <- list(5, 10, -3)
sigma <- list(1, 5, 10)
n <- list(10, 30, 50)
args1 <- list(n, mu, sigma)

pmap(args1, rnorm)

args1 %>%
  pmap(rnorm) %>% 
  str()


```
## Pass values through tribble to pmap
```{r}
params <- tribble(
  ~mean, ~sd, ~n,
    5,     1,  1,
   10,     5,  3,
   -3,    10,  5
)

pmap(params, rnorm)
#or
params %>% 
  pmap(rnorm)
```
## Invoking different functions
```{r}
f <- c("runif", "rnorm", "rpois")

param <- list(
  list(min = -1, max = 1), 
  list(sd = 5), 
  list(lambda = 10)
)

invoke_map(f, param, n=5)

invoke_map(f, param, n = 5) %>% glimpse()
```


## Reference
* https://r4ds.had.co.nz/iteration.html