---
title: "Mid-Sem Solutions"
author: "PfA Teaching Team"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# packages required
library(readxl)
library(tidyverse)
#library(janitor)
# Doing this without using janitor package

# remove scientific notation
options(scipen=999)
```


## Q.No. 1
## Read the excel file "Indian Startup Dataset.xlsx" into a startup_funding variable as dataframe
* Read the dataset to R environment
```{r}
startup_funding <- read_excel("G:/My Drive/Deep Learning Workshop/Indian Startup Dataset.xlsx",
skip = 2) 
#startup_funding
head(startup_funding)
```
## Q.No. 2
## Clean the column names to small cases, remove special characters, replace space with underscore

```{r}
colnames(startup_funding)
```

```{r}
startup_funding <- startup_funding %>% rename( sr_no = `Sr. No.`,
                            date = `Date (dd/mm/yyyy)`,
                            startup_name = `Startup Name`,
                            industry_vertical = `Industry / Vertical`,
                            sub_vertical = `Sub-Vertical`,
                            city_location = `City / Location`,
                            investors_name = `Investors' Name`,
                            investment_type = `Investment Type`,
                            amount_usd = `Amount (In USD)`
                            )

colnames(startup_funding)
```
## Q.No. 3
## Create a column containing unique value per row 
```{r}
dim(startup_funding)
```
```{r}
#nrow() provides number if dimensions
startup_funding$row_id <- seq(1, nrow(startup_funding),1)

#rows with missing have a common pattern with date
startup_funding <- startup_funding %>% 
  filter(date != "00/01/1900") %>% 
  select(row_id, everything())

startup_funding
```

```{r}
startup_funding <- startup_funding %>% 
  mutate(amount_cleaned = str_replace_all(amount_usd, ",", ""))
```


```{r}
startup_funding <- startup_funding %>% 
  mutate(amount_cleaned = ifelse(amount_cleaned %in% c("N/A", "NA" , "Undisclosed" , "undisclosed" , "unknown" , "upto 15000000"), "undisclosed", as.numeric(amount_cleaned))
         ) 

head(startup_funding)


```

```{r}
startup_funding_undisclosed <-  startup_funding %>% filter(amount_cleaned == "undisclosed") 

startup_funding_undisclosed
```

```{r}
startup_funding <-  startup_funding %>% filter(amount_cleaned != "undisclosed") %>% 
  mutate(amount_cleaned = as.numeric(amount_cleaned),
         amount_cleaned_inr = amount_cleaned*75
         )
```

## Q.No. 5
## "Create separate columns
* (a) day
* (b) month
* (c ) year
* Extract day, month and year from date column, and populate the same in (a), (b), and (c )


```{r}
startup_funding %>% 
  mutate(date_new = lubridate::dmy(date),
         day      = lubridate::day(date_new),
         month    = lubridate::month(date_new),
         year     = lubridate::year(date_new)
         )  %>% 
  select(date,  date_new, day, month, year)

#or

startup_funding <- startup_funding %>% 
  mutate(date_new = lubridate::dmy(date),
         day      = as.numeric(str_sub(date_new,9,10)),
         month    = as.numeric(str_sub(date_new,6,7)),
         year     = as.numeric(str_sub(date_new,1,4))
         ) 

```

## Q.No. 6
## Identify the top 10 highest:
* (1) funded startups
* (2) Industry Verticals
* (3) Industry Verticals & sub-verticals
* (4) Location
* (5) type of investment

```{r}
startup_funding %>% head()
```


```{r}
#by startup_name
startup_funding %>% group_by(startup_name) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(desc(total_funding_inr)) %>% 
  select(startup_name, total_funding_inr) %>% head(10)

#by industry_vertical
startup_funding %>% group_by(industry_vertical) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(desc(total_funding_inr)) %>% 
  select(industry_vertical, total_funding_inr) %>% head(10)


#by industry_vertical & sub_vertical
startup_funding %>% group_by(industry_vertical, sub_vertical) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(desc(total_funding_inr)) %>% 
  select(industry_vertical,  sub_vertical, total_funding_inr) %>% head(10)

#by location
startup_funding %>% group_by(city_location) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(desc(total_funding_inr)) %>% 
  select(city_location, total_funding_inr) %>% head(10)

# by type of investment
startup_funding %>% group_by(investment_type) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(desc(total_funding_inr)) %>% 
  select(investment_type, total_funding_inr) %>% head(10)


```


## Q.No. 7
## Provide the summary statistics of funded startups by industry vertical, sub-vertical, type of investment and location


```{r}
#industry_vertical, sub_vertical, city_location, investment_type
startup_funding %>% 
  group_by(industry_vertical, sub_vertical, city_location, investment_type ) %>% 
  summarise( count_all  = n(),
            sum_al = sum(amount_cleaned_inr),
            mean_al = mean(amount_cleaned_inr),
            median_al = median(amount_cleaned_inr),
            sd_al     =  sd(amount_cleaned_inr)
            ) %>% 
  arrange(desc(count_all))


```


## Q.No. 8
## Explore the following for the startups with undisclosed amount funded: 
 * (a) Identify which of them were funded with disclosed amount previously and after
 * (b) Identify which of them were not funded 
 
```{r}
startup_funding_undisclosed
```

```{r}
startup_funding
```

```{r}
startup_funding_undisclosed %>% inner_join(startup_funding, by = "startup_name")

# Not funded previously or after
startup_funding_undisclosed %>% anti_join(startup_funding, by = "startup_name")




```

## Q.No. 9
## By year, identify the investment type, and amount per startup. Display the results in descending order of year.

```{r}
startup_funding %>% group_by(startup_name, year, investment_type ) %>% 
  summarise(total_funding_inr = sum(amount_cleaned_inr)) %>% 
  arrange(startup_name, desc(year)) %>% 
  select(startup_name, year, investment_type, total_funding_inr)
```

## Q.No. 10
## Find the startups funded with 
* (a) more than 50 million USD, or 400 Crores INR 
* (b) that are not Fintech, Saas, Transportation, Finance, and E-Commerce
* (c) provide total funding of these startups


```{r}
startup_funding %>% 
  filter(!industry_vertical %in% c("Fintech", "Saas", "Finance", "Transportation", "E-Commerce")) %>% 
  group_by(startup_name) %>% 
  summarise( total_funding_usd = sum(amount_cleaned),
    total_funding_inr = sum(amount_cleaned_inr)) %>% 
  filter(total_funding_usd > 50000000, total_funding_inr > 4000000000) %>% 
  select(startup_name, total_funding_usd, total_funding_inr)
```

