---
title: "Categorical Data (Operations)"
author: "Revendranath T"
date: "6/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Categorical Data Types
1. Nominal 
  * Unordered data such as Gender (Male, Female, Binary)
2. Ordinal
  * Ordered data such as ranks of students in a class. Another example, Likert scale (Strongly disagree to strongly agree)

## Attach packages
```{r}
library(tidyverse)
library(janitor)
```


## use factor in R to represent nominal data
```{r}
#Review the difference
gender <- c("male", "female", "binary")
gender
typeof(gender)
class(gender)
levels(gender)
```


```{r}
gender_factor <- factor(c("male", "female", "binary"))
gender_factor
typeof(gender_factor)
class(gender_factor)
levels(gender_factor)
nlevels(gender_factor)
min(gender_factor) # no minimum values in nominal data (not-ordered, not ordinal)

gender_factor <- factor(rep(c("male", "female", "male", "binary"), 3))
gender_factor
summary(gender_factor)
```

#Ordered factors or ordinal data
```{r}
difficulty <- factor(c("easy", "medium", "hard"), ordered = TRUE)
difficulty
difficulty <-  fct_inorder(difficulty)

difficulty
typeof(difficulty)
class(difficulty)
levels(difficulty)
nlevels(difficulty)
min(difficulty) 
max(difficulty)

library(forcats)
difficulty <-  fct_relevel(difficulty, c("easy", 
                          "medium", 
                          "hard"))

difficulty
```
##  Get serious!
```{r}
library(RCurl)
# x <- getURL("https://raw.githubusercontent.com/Defcon27/Data-Analysis-of-Indian-Automobile-dataset-using-Machine-Learning-in-R/master/indian-auto-mpg.csv"
#             )
# 
# impg <- read.csv(text = x)

#library(janitor)
impg <- read.csv("https://raw.githubusercontent.com/Defcon27/Data-Analysis-of-Indian-Automobile-dataset-using-Machine-Learning-in-R/master/indian-auto-mpg.csv") %>% clean_names()

impg %>% glimpse()

impg %>% select(name) %>%  unique() %>% get_dupes()

impg %>% select(manufacturer) %>%  unique() %>% get_dupes()

impg %>% select(transmission)  %>%  unique() %>% get_dupes()

impg %>% select(fuel_type)  %>%  unique() %>% get_dupes()

impg %>% select(owner_type)  %>%  unique() %>% get_dupes()

impg %>% select(engine_cc)  %>%  unique() %>% get_dupes()

impg %>% select(power)  %>%  unique() %>% get_dupes()

impg %>% select(seats)  %>%  unique() %>% get_dupes()

impg %>% select(mileage_km_l)  %>%  unique() %>% get_dupes()
```


```{r}
library(forcats)
impg %>% glimpse()



#convert to factors
impg$manufacturer <- as.factor(impg$manufacturer)
impg$fuel_type <- as.factor(impg$fuel_type)
impg$transmission <- as.factor(impg$transmission)
impg$owner_type <- as.factor(impg$owner_type)
impg$location <- as.factor(impg$location)          

#another method to convert to factors using map
impg <- impg %>% 
  mutate(manufacturer = as.factor(manufacturer), 
         fuel_type = as.factor(fuel_type), 
         transmission = as.factor(transmission), 
         owner_type = as.factor(owner_type), 
         location = as.factor(location)
         ) 

impg %>% glimpse()

impg %>% 
  select(manufacturer, fuel_type, transmission, owner_type) %>% 
  map(levels)
#or
levels(impg$manufacturer)
levels(impg$fuel_type) 
levels(impg$transmission) 
levels(impg$owner_type) 
levels(impg$location) 

#Count the number of values: fct_count from forecats package
fct_count(impg$manufacturer)
fct_count(impg$manufacturer, sort = TRUE) 
fct_count(impg$fuel_type, sort = TRUE) 
fct_count(impg$transmission, sort = TRUE) 
fct_count(impg$owner_type, sort = TRUE) 

# impg %>% 
#   select(manufacturer, fuel_type, transmission, owner_type) %>% 
#   summarise(manu_cnt = fct_count(manufacturer, sort=TRUE)) %>% 
#   unnest_longer(col = c(manu_cnt)) # --> Throws error while unnest-ing

#Remove duplicates to get unique values
fct_unique(impg$manufacturer) 
fct_unique(impg$fuel_type) 
fct_unique(impg$transmission) 
fct_unique(impg$owner_type) 
fct_unique(impg$location) 


# Count and unique pose challenges due to NAs
x <- factor(c("a", "b", NA, "c"))
x
fct_unique(x)
fct_count(x)
```
## Get the owner-type count and sort them by frequency of owner-type category
```{r}
fct_count(impg$owner_type, sort = TRUE) 
```

## Changing the order of levels for a factor variable

### Manually reordering factor variables
```{r}
difficulty <- factor(c("easy", "medium", "hard"), ordered = TRUE)
difficulty

difficulty_unordered <- factor(c("medium", "easy", "very easy", "hard"), 
                               ordered = TRUE)
difficulty_unordered

difficulty_ordered <- fct_relevel(difficulty_unordered, 
                                  c("very easy", 
                                    "easy", 
                                    "medium", 
                                    "hard")) 

difficulty_ordered
```
## Reorder factor levels based on the appearance in data
```{r}

difficulty_unordered_1 <- rep(c("medium", "easy", "very easy", "hard"),3)

difficulty_unordered_1 <- factor(difficulty_unordered_1, ordered = TRUE)

difficulty_unordered_1

#Reorder factor levels based on the appearance in data

fct_inorder(difficulty_unordered_1)
#factors are not ordered. Therefore, reorder

difficulty_unordered_1 <- fct_relevel(difficulty_unordered_1, 
                                  c("very easy", 
                                    "easy", 
                                    "medium", 
                                    "hard")) 

difficulty_unordered_1

# Even then, factor levels are order based on the appearance in data
difficulty_unordered_1 <- fct_inorder(difficulty_unordered_1)

difficulty_unordered_1

```
## Order factor levels based on the frequency
```{r}

difficulty_unordered_2 <- rep(c("medium", "easy", "medium", "very easy", "hard"),3)
difficulty_unordered_2
difficulty_unordered_2 <- factor(difficulty_unordered_2)
difficulty_unordered_2

fct_infreq(difficulty_unordered_2)


difficulty_unordered_3 <- factor(difficulty_unordered_2, ordered = TRUE)

difficulty_unordered_3 <- fct_infreq(difficulty_unordered_3)

x <- c(rep("medium",3), rep("easy",2), rep("very easy",7), rep("hard",11))
x
x <- factor(x)
x
fct_infreq(x)
```

## Reversing the order of levels using fct_rev
```{r}
print(levels(difficulty_unordered_3))

print(levels(fct_rev(difficulty_unordered_3)))

```

## Reorder factor levels based on the relationship with other variables
## use fct_reorder
```{r}
df <- tibble::tribble(
  ~color,     ~a, ~b,
  "blue",      1,  2,
  "green",     6,  2,
  "purple",    3,  3,
  "red",       2,  3,
  "yellow",    5,  1
)
df
df$color <- factor(df$color)

df

fct_reorder(df$color, df$a, min)

print(fct_reorder(df$color, df$b, min))
#print(fct_reorder(df$color, df$b, max)) (Review & Refer to documentation of forcats)
```
## Add or drop factor levels in R
The three functions which are important to know from the addition and deletion perspective are
* fct_expand() – use it to add new level
* fct_explicit_na() – use it, if you wish to assign NA as one of the levels. This way, when you plot charts, NA’s will also appear.
* fct_drop() – use it drop a particular level
```{r}

# Adding factor level
difficulty <- fct_expand(difficulty,"very very easy")
difficulty

# Converting NA to factor level
f1 <- factor(c(1, 1, NA, NA,2, 2, NA,2, 1, 2, 2))
f2 <- fct_explicit_na(f1, na_level = "(Unknown)")
print(f1)
print(f2)

fct_unique(f1)
fct_count(f1)

fct_unique(f2)
fct_count(f2)

# Drop factor level --> (The documentation is limited to provide illustration)
fac1 <- factor(c(c("aa","bb"), c("aa","bb","cc")))
fac1
fac2 <- fct_drop(fac1) 
fac2
fac1

print(fac1)
print(fac2)


x <- c(c("aa","bb"), c("aa","bb","cc"))
x
length(x)


#Reference from: https://forcats.tidyverse.org/reference/fct_drop.html
f <- factor(c("a", "b"), levels = c("a", "b", "c"))
f
fct_drop(f)
# Set only to restrict which levels to drop
fct_drop(f, only = "a")

fct_drop(f, only = "c")

```
## Changing values of factor levels in R
* Manually combine levels into defined groups
```{r}
#collapse using fct_collapse()
difficulty_easy <-  fct_expand(difficulty,c("very very easy", "very easy"))
difficulty_easy

df1 <- fct_collapse(difficulty_easy, easy = c("easy", "very easy", "very very easy"))
df1
```

## Show levels to others or hide them
```{r}
difficulty <- factor(c("easy", "medium", "hard"), ordered = TRUE)
difficulty
# Example showing keep as argument to keep a level, and hide other levels
fct_other(difficulty, keep = c("easy"))
# Example showing drop as argument to hide a level, and keep other levels visible
fct_other(difficulty, drop = c("easy"))

fct_other(difficulty, keep = c("hard", "medium"))
```
## Shuffle and Shift the levels
```{r}
diff <- factor(c(rep("medium",3), 
                 rep("easy",2), 
                 rep("very easy",4), 
                 rep("hard",5)))
diff 

fct_lump(diff, n = 1)
print(diff)
print(fct_lump(diff, n = 2))
diff

#Shuffle the levels
print(diff)
print(fct_shuffle(diff))

#Shift the levels
print(diff)
print(fct_shift(diff,2))
```

## Group most/least common levels into a single level
```{r}
diff <- factor(c(rep("medium",3), rep("easy",2), rep("very easy",4), rep("hard",5)))
diff 

fct_lump(diff, n = 1)

x <- factor(rep(LETTERS[1:9], times = c(40, 10, 5, 27, 1, 1, 1, 1, 1)))
x %>% table()
x %>% fct_lump_n(3) %>% table()
x %>% fct_lump_prop(0.10) %>% table()
x %>% fct_lump_min(5) %>% table()
x %>% fct_lump_lowfreq() %>% table()

```
```{r}
x <- c("a", "z", "g")
x <- c("easy", "medium", "hard")
as.factor(x)

as_factor(x)

factor(x)
```


## References
* https://r4ds.had.co.nz/factors.html
* https://www.r-bloggers.com/2020/06/working-with-factors-in-r-tutorial-forcats-package/
* https://forcats.tidyverse.org/
